---
title: "Final_TensRule"
author: "Will Pfadenhauer"
date: "2022-12-22"
output: html_document
---

Load data & packages
```{r}
require(data.table)
require(dplyr)
require(tidyr)
require(naniar)
require(ggplot2)
require(Hmisc)

#Load data
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")

#data_L1_I <- data[!(data$Invasive_L1 == ""), ]
#data_L3_I <- data[!(data$Invasive_L3 == ""), ]
#New_L2_Invaders <- anti_join(data_L1_I, data_L3_I, by="Accepted_name")
#data <- anti_join(data, New_L2_Invaders, by="Accepted_name")
```


**WGSRPD Level 1 - Continents**

Make a Data Table to Store continent Results
```{r}
AllContsDestinations <- data.table()
AllContsDestinations$Continent <- c(",1,", ",2,", ",3,", ",4,", ",5,", ",6,", ",7,", ",8,", ",9,")
AllContsDestinations$Names <- c("Europe", "Africa", "Asia Temperate", "Asia Tropical", "Australasia", "Pacific", "Northern America", "Southern America", "Antarctic")
AllContsDestinations$EstablishedSpecies <- ""
AllContsDestinations$InvasiveSpecies <- ""
```

Count established and invasive species for each continent
```{r}
a <- c(1:9)
for (i in a) {
  
#Need species with Invasive_L1 = whatever L1 region is in Row i
Region_I <- data[grep(paste(AllContsDestinations[i,1] ), data$Invasive_L1),]

#Remove species with Native L1 = whatever L1 region is in Row i 
Region_I <- Region_I[!(grep(paste(AllContsDestinations[i,1]), Region_I$Native_L1)), ]

#Remove excess columns
Region_I <- Region_I[,c(1)]

#Label as invasive
Region_I$type = "1"

#Need species with Established_L1 = whatever L1 region is in Row i
Region_E <- data[grep(paste(AllContsDestinations[i,1] ), data$Established_L1),]

#Remove species with Native L1 = whatever L1 region is in Row i
Region_E <- Region_E[!(grep(paste(AllContsDestinations[i,1]), Region_E$Native_L1)),]

#Remove invasive species from our established list
Region_E <-  Region_E[!(grep(paste(AllContsDestinations[i,1] ), Region_E$Invasive_L1)), ]

#Remove excess columns
Region_E <- Region_E[,c(1)]

#Label as established
Region_E$type = "0"

#Rbind Invaders to Established 
Region <- rbind(Region_I, Region_E)

Region_count <- table(Region$type)

AllContsDestinations[i,3] <- paste(Region_count[1])
AllContsDestinations[i,4] <- paste(Region_count[2])

rm(Region)
rm(Region_E)
rm(Region_I)
rm(Region_count)
  
 message('Running Region ', i, ' of 9')

}

#Calculate Invasion Rates
AllContsDestinations$EstablishedSpecies <- as.integer(AllContsDestinations$EstablishedSpecies)
AllContsDestinations$InvasiveSpecies <- as.integer(AllContsDestinations$InvasiveSpecies)
AllContsDestinations$Total <- (AllContsDestinations$EstablishedSpecies + AllContsDestinations$InvasiveSpecies)

#Only use this line of code if you're doing the version of the analysis presented in Table S5. 
#AllContsDestinations <- subset(AllContsDestinations, Total >= 50)

AllContsDestinations$InvasionRate <- (AllContsDestinations$InvasiveSpecies/AllContsDestinations$Total)

#What's the average invasion rate for all continents?
#This is different than what's reported in the manuscript because this excludes Antarctica. 
mean(AllContsDestinations$InvasionRate)

#Prepare data for plotting
pd_1 <- AllContsDestinations[,c(1,6)]
pd_1$Scale <- "L1"

#Remove commas before exporting
AllContsDestinations$Continent <- gsub(",", "", AllContsDestinations$Continent)

#Export for mapping in ArcGIS Pro
#fwrite(AllContsDestinations, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/ContsMappingData.csv")
```

Compare L1 average to Tens Rule (Significance Testing)
```{r}
test <- t.test(AllContsDestinations$InvasionRate,
  mu = 0.10)
test
#T-test says L1 average is significantly greater than 10%.

test1 <- t.test(AllContsDestinations$InvasionRate,
  mu = 0.20)
test1
#T-test says L1 average is significantly different than 20%.

#Wilcoxon test
test1mw <- wilcox.test(AllContsDestinations$InvasionRate, alternative = "greater", mu = 0.10)
test1mw
#Even one-sided, nonparametric test agrees that L1 average is significantly greater than 10%.

test1tr <- wilcox.test(AllContsDestinations$InvasionRate, alternative = "greater", mu = 0.20)
test1tr
#Even one-sided, nonparametric test agrees that L1 average is significantly greater than 20%.

#Binomial GLM
AllContsDestinations$os <-qlogis(0.1)
test_glm <- glm(cbind(AllContsDestinations$InvasiveSpecies,AllContsDestinations$EstablishedSpecies) ~ offset(AllContsDestinations$os) + 1, family = binomial(link="logit"))

summary(test_glm)
#GLM agrees with previous two tests
```
I recommend removing all objects except pd_1 from workspace before proceeding to L2 regions. 

**WGSRPD Level 2 - Sub-Continental Regions**

Need to create a list of all L2 regions
```{r}
#Reload data just to be safe
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")

#data_L1_I <- data[!(data$Invasive_L1 == ""), ]
#data_L3_I <- data[!(data$Invasive_L3 == ""), ]
#New_L2_Invaders <- anti_join(data_L1_I, data_L3_I, by="Accepted_name")
#data <- anti_join(data, New_L2_Invaders, by="Accepted_name")

#Expand L4 regions
data_e <- separate_rows(data, Established_L2, sep = ",")
data_i <- separate_rows(data, Invasive_L2, sep = ",")
data_n <- separate_rows(data, Native_L2, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,8)]
data_i <- data_i[,c(1,12)]
data_n <- data_n[,c(1,4)]

#rename for matching
data_e <- rename(data_e, "L2" ="Established_L2")
data_i <- rename(data_i, "L2" ="Invasive_L2")
data_n <- rename(data_n, "L2" ="Native_L2")

#Combine all L4 regions
full_l2 <-rbind(data_e, data_i)
full_l2 <- rbind(full_l2, data_n)

#Remove duplicates
full_l2 <- distinct(full_l2,
                    L2)

#Sort by numeric order
full_l2$L2 <- as.integer(full_l2$L2)
full_l2 <- full_l2[order(full_l2$L2),]

#Remove NA row
full_l2 <- full_l2[-c(53),]
```

Make a Data Table to Store L2 Results
```{r}
All2Destinations <- data.table()
All2Destinations$L2s <- paste(",", full_l2$L2, ",", sep = "")
All2Destinations$EstablishedSpecies <- ""
All2Destinations$InvasiveSpecies <- ""
```

Count established and invasive species for each L2 region
```{r}
a <- c(1:52)
for (i in a) {
  
#Need species with Invasive_L2 = whatever L2 region is in Row i
Region_I <- data[grep(paste(All2Destinations[i,1] ), data$Invasive_L2),]

#Remove species with Native L2 = whatever L2 region is in Row i 
Region_I <- Region_I[!(grep(paste(All2Destinations[i,1]), Region_I$Native_L2)), ]

#Remove excess columns
Region_I <- Region_I[,c(1)]

#Label as invasive
Region_I$type = "1"

#Need species with Established_L2 = whatever L2 region is in Row i
Region_E <- data[grep(paste(All2Destinations[i,1] ), data$Established_L2),]

#Remove species with Native L2 = whatever L2 region is in Row i
Region_E <- Region_E[!(grep(paste(All2Destinations[i,1]), Region_E$Native_L2)),]

#Remove invasive species from our established list
Region_E <-  Region_E[!(grep(paste(All2Destinations[i,1] ), Region_E$Invasive_L2)), ]

#Remove excess columns
Region_E <- Region_E[,c(1)]

#Label as established
Region_E$type = "0"

#Rbind Invaders to Established 
Region <- rbind(Region_I, Region_E)

Region_count <- table(Region$type)

All2Destinations[i,2] <- paste(Region_count[1])
All2Destinations[i,3] <- paste(Region_count[2])

rm(Region)
rm(Region_E)
rm(Region_I)
rm(Region_count)
  
 message('Running Region ', i, ' of 52')

}
```

```{r}
#Remove NAs
All2Destinations <- as.data.frame(All2Destinations)
All2Destinations <- replace_with_na(All2Destinations, replace = list(EstablishedSpecies = c("NA")))
All2Destinations <- replace_with_na(All2Destinations, replace = list(InvasiveSpecies = c("NA")))
L2NAs <- All2Destinations[!complete.cases(All2Destinations), ]
All2Destinations <- na.omit(All2Destinations)

#Calculate Invasion Rates
All2Destinations$EstablishedSpecies <- as.integer(All2Destinations$EstablishedSpecies)
All2Destinations$InvasiveSpecies <- as.integer(All2Destinations$InvasiveSpecies)
All2Destinations$Total <- (All2Destinations$EstablishedSpecies + All2Destinations$InvasiveSpecies)

#Only use this line of code if you're doing the version of the analysis presented in Table S5. 
#All2Destinations <- subset(All2Destinations, Total >= 50)

All2Destinations$InvasionRate <- (All2Destinations$InvasiveSpecies/All2Destinations$Total)

#What's the average invasion rate for all L2 regions?
mean(All2Destinations$InvasionRate)

#Prepare data for plotting
pd_2 <- All2Destinations[,c(1,5)]
pd_2$Scale <- "L2"

#Remove commas before exporting
All2Destinations$L2s <- gsub(",", "", All2Destinations$L2s)

#Export for mapping in ArcGIS Pro
#fwrite(All2Destinations, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/L2MappingData.csv")
```

Compare L2 average to Tens Rule (Significance Testing)
```{r}
test2 <- t.test(All2Destinations$InvasionRate,
  mu = 0.10)
test2
#T-test says L2 average is significantly greater than 10%.

test22 <- t.test(All2Destinations$InvasionRate,
  mu = 0.20)
test22
#T-test says L2 average is not significantly different than 20%.

#Wilcoxon test
test2mw <- wilcox.test(All2Destinations$InvasionRate, alternative = "greater", mu = 0.10)
test2mw
#Even one-sided, nonparametric test agrees that L2 average is significantly greater than 10%.

test2tr <- wilcox.test(All2Destinations$InvasionRate, mu = 0.20)
test2tr
#Nonparametric test agrees that L2 average is not significantly different than 20%.

#Binomial GLM
All2Destinations$os <-qlogis(0.1)
test_glm2 <- glm(cbind(All2Destinations$InvasiveSpecies,All2Destinations$EstablishedSpecies) ~ offset(All2Destinations$os) + 1, family = binomial(link="logit"))

summary(test_glm2)
#GLM agrees with previous two tests
```
I recommend removing all objects except pd_1 and pd_2 from workspace before proceeding to L3 regions. 

**WGSRPD Level 3 - Countries/States**

Need to create a list of all L3 regions
```{r}
#Reload data just to be safe
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")

#data_L1_I <- data[!(data$Invasive_L1 == ""), ]
#data_L3_I <- data[!(data$Invasive_L3 == ""), ]
#New_L2_Invaders <- anti_join(data_L1_I, data_L3_I, by="Accepted_name")
#data <- anti_join(data, New_L2_Invaders, by="Accepted_name")

#Expand L3 regions
data_e <- separate_rows(data, Established_L3, sep = ",")
data_i <- separate_rows(data, Invasive_L3, sep = ",")
data_n <- separate_rows(data, Native_L3, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,9)]
data_i <- data_i[,c(1,13)]
data_n <- data_n[,c(1,5)]

#rename for matching
data_e <- rename(data_e, "L3" ="Established_L3")
data_i <- rename(data_i, "L3" ="Invasive_L3")
data_n <- rename(data_n, "L3" ="Native_L3")

#Combine all L3 regions
full_l3 <-rbind(data_e, data_i)
full_l3 <- rbind(full_l3, data_n)

#Remove duplicates
full_l3 <- distinct(full_l3,
                    L3)

#Sort by alphabetical order
full_l3 <- full_l3[order(full_l3$L3),]

#Remove weird rows
full_l3 <- full_l3[-c(1),]
```

Make a Data Table to Store L3 Results
```{r}
All3Destinations <- data.table()
All3Destinations$L3s <- paste(",", full_l3$L3, ",", sep = "")
All3Destinations$EstablishedSpecies <- ""
All3Destinations$InvasiveSpecies <- ""
```

Count established and invasive species for each L3 region
```{r}
a <- c(1:368)
for (i in a) {
  
#Need species with Invasive_L3 = whatever L3 region is in Row i
Region_I <- data[grep(paste(All3Destinations[i,1] ), data$Invasive_L3),]

#Remove species with Native L3 = whatever L3 region is in Row i 
Region_I <- Region_I[!(grep(paste(All3Destinations[i,1]), Region_I$Native_L3)), ]

#Remove excess columns
Region_I <- Region_I[,c(1)]

#Label as invasive
Region_I$type = "1"

#Need species with Established_L3 = whatever L3 region is in Row i
Region_E <- data[grep(paste(All3Destinations[i,1] ), data$Established_L3),]

#Remove species with Native L3 = whatever L3 region is in Row i
Region_E <- Region_E[!(grep(paste(All3Destinations[i,1]), Region_E$Native_L3)),]

#Remove invasive species from our established list
Region_E <-  Region_E[!(grep(paste(All3Destinations[i,1] ), Region_E$Invasive_L3)), ]

#Remove excess columns
Region_E <- Region_E[,c(1)]

#Label as established
Region_E$type = "0"

#Rbind Invaders to Established 
Region <- rbind(Region_I, Region_E)

Region_count <- table(Region$type)

All3Destinations[i,2] <- paste(Region_count[1])
All3Destinations[i,3] <- paste(Region_count[2])

rm(Region)
rm(Region_E)
rm(Region_I)
rm(Region_count)
  
 message('Running Region ', i, ' of 368')

}
```

```{r}
#Remove NAs
All3Destinations <- as.data.frame(All3Destinations)
All3Destinations <- replace_with_na(All3Destinations, replace = list(EstablishedSpecies = c("NA")))
All3Destinations <- replace_with_na(All3Destinations, replace = list(InvasiveSpecies = c("NA")))
L3NAs <- All3Destinations[!complete.cases(All3Destinations), ]
All3Destinations <- na.omit(All3Destinations)

#Calculate Invasion Rates
All3Destinations$EstablishedSpecies <- as.integer(All3Destinations$EstablishedSpecies)
All3Destinations$InvasiveSpecies <- as.integer(All3Destinations$InvasiveSpecies)
All3Destinations$Total <- (All3Destinations$EstablishedSpecies + All3Destinations$InvasiveSpecies)

#Only use this line of code if you're doing the version of the analysis presented in Table S5. 
#All3Destinations <- subset(All3Destinations, Total >= 50)

All3Destinations$InvasionRate <- (All3Destinations$InvasiveSpecies/All3Destinations$Total)

#What's the average invasion rate for all L3 regions?
mean(All3Destinations$InvasionRate)

#Prepare data for plotting
pd_3 <- All3Destinations[,c(1,5)]
pd_3$Scale <- "L3"

#Remove commas before exporting
All3Destinations$L3s <- gsub(",", "", All3Destinations$L3s)

#Export for mapping in ArcGIS Pro
#fwrite(All3Destinations, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/L3MappingData.csv")
```


Compare L3 average to Tens Rule (Significance Testing)
```{r}
test3 <- t.test(All3Destinations$InvasionRate, mu = 0.10)
test3
#T-test says L3 average is significantly different than 10%.

test33 <- t.test(All3Destinations$InvasionRate, alternative = "less", mu = 0.20)
test33
#T-test says L3 average is significantly less than 20%.


#Wilcoxon test
test3mw <- wilcox.test(All3Destinations$InvasionRate, alternative = "greater", mu = 0.10)
test3mw
#Even one-sided, nonparametric test agrees that L3 average is significantly greater than 10%.

test3tr <- wilcox.test(All3Destinations$InvasionRate, alternative = "less", mu = 0.20)
test3tr
#Nonparametric test agrees that L3 values are significantly lower than 20%. 

#Binomial GLM
All3Destinations$os <-qlogis(0.1)
test_glm3 <- glm(cbind(All3Destinations$InvasiveSpecies,All3Destinations$EstablishedSpecies) ~ offset(All3Destinations$os) + 1, family = binomial(link="logit"))

summary(test_glm3)
#GLM agrees with previous two tests

```
I recommend removing all objects except pd_1, pd_2, and pd_3 from workspace and reloading the first code chunk at the top before proceeding to L4 regions. 


**WGSRPD Level 4 - Small Countries/ States**

Need to create a list of all L4 regions
```{r}
#Reload data just to be safe
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")

#data_L1_I <- data[!(data$Invasive_L1 == ""), ]
#data_L3_I <- data[!(data$Invasive_L3 == ""), ]
#New_L2_Invaders <- anti_join(data_L1_I, data_L3_I, by="Accepted_name")
#ata <- anti_join(data, New_L2_Invaders, by="Accepted_name")

#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,10)]
data_i <- data_i[,c(1,14)]
data_n <- data_n[,c(1,6)]

#rename for matching
data_e <- rename(data_e, "L4" ="Established_L4")
data_i <- rename(data_i, "L4" ="Invasive_L4")
data_n <- rename(data_n, "L4" ="Native_L4")

#Combine all L4 regions
full_l4 <-rbind(data_e, data_i)
full_l4 <- rbind(full_l4, data_n)

#Remove duplicates
full_l4 <- distinct(full_l4,
                    L4)

#Remove regions ending in -OO (not true L4 regions - those are converted L3 regions)
full_l4 <- as.data.frame(full_l4[!grepl("-OO", full_l4$L4),])

#Sort by alphabetical order
full_l4 <- as.data.table(full_l4[order(full_l4$L4),])

#Remove weird rows
full_l4 <- as.data.table(full_l4[-c(1),])
```

Make a Data Table to Store L4 Results
```{r}
All4Destinations <- data.table()
All4Destinations$L4s <- paste(",", full_l4$V1, ",", sep = "")
All4Destinations$EstablishedSpecies <- ""
All4Destinations$InvasiveSpecies <- ""
```

Count established and invasive species for each L4 region
```{r}
a <- c(1:271)
for (i in a) {
  
#Need species with Invasive_L4 = whatever L4 region is in Row i
Region_I <- data[grep(paste(All4Destinations[i,1] ), data$Invasive_L4),]

#Remove species with Native L4 = whatever L4 region is in Row i 
Region_I <- Region_I[!(grep(paste(All4Destinations[i,1]), Region_I$Native_L4)), ]

#Remove excess columns
Region_I <- Region_I[,c(1)]

#Label as invasive
Region_I$type = "1"

#Need species with Established_L4 = whatever L4 region is in Row i
Region_E <- data[grep(paste(All4Destinations[i,1] ), data$Established_L4),]

#Remove species with Native L4 = whatever L4 region is in Row i
Region_E <- Region_E[!(grep(paste(All4Destinations[i,1]), Region_E$Native_L4)),]

#Remove invasive species from our established list
Region_E <-  Region_E[!(grep(paste(All4Destinations[i,1] ), Region_E$Invasive_L4)), ]

#Remove excess columns
Region_E <- Region_E[,c(1)]

#Label as established
Region_E$type = "0"

#Rbind Invaders to Established 
Region <- rbind(Region_I, Region_E)

Region_count <- table(Region$type)

All4Destinations[i,2] <- paste(Region_count[1])
All4Destinations[i,3] <- paste(Region_count[2])

rm(Region)
rm(Region_E)
rm(Region_I)
rm(Region_count)
  
 message('Running Region ', i, ' of 271')

}
```

```{r}
#Remove NAs
All4Destinations <- as.data.frame(All4Destinations)
All4Destinations <- replace_with_na(All4Destinations, replace = list(EstablishedSpecies = c("NA")))
All4Destinations <- replace_with_na(All4Destinations, replace = list(InvasiveSpecies = c("NA")))
L4NAs <- All4Destinations[!complete.cases(All4Destinations), ]
All4Destinations <- na.omit(All4Destinations)

#Calculate Invasion Rates
All4Destinations$EstablishedSpecies <- as.integer(All4Destinations$EstablishedSpecies)
All4Destinations$InvasiveSpecies <- as.integer(All4Destinations$InvasiveSpecies)
All4Destinations$Total <- (All4Destinations$EstablishedSpecies + All4Destinations$InvasiveSpecies)

#Only use this line of code if you're doing the version of the analysis presented in Table S5. 
#All4Destinations <- subset(All4Destinations, Total >= 50)

All4Destinations$InvasionRate <- (All4Destinations$InvasiveSpecies/All4Destinations$Total)

#What's the average invasion rate for all L4 regions?
mean(All4Destinations$InvasionRate)

#Prepare data for plotting
pd_4 <- All4Destinations[,c(1,5)]
pd_4$Scale <- "L4"

#Remove commas before exporting
All4Destinations$L4s <- gsub(",", "", All4Destinations$L4s)

#Export for mapping in ArcGIS Pro
#fwrite(All4Destinations, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/L4MappingData.csv")
```

Compare L4 average to Tens Rule (Significance Testing)
```{r}
test4 <- t.test(All4Destinations$InvasionRate, mu = 0.10)
test4
#T-test says L4 average is not significantly different than 10%.

#Wilcoxon test
test4mw <- wilcox.test(All4Destinations$InvasionRate, alternative = "less", mu = 0.10)
test4mw
# This is the only time the nonparametric test disagrees with the t-test.
# Nonparametric test thinks L4 average is significantly less than 10%. 

#Binomial GLM
All4Destinations$os <-qlogis(0.1)
test_glm4 <- glm(cbind(All4Destinations$InvasiveSpecies,All4Destinations$EstablishedSpecies) ~ offset(All4Destinations$os) + 1, family = binomial(link="logit"))

summary(test_glm4)
#GLM agrees with t-test, but not Wilcoxon. 
```

**Confidence Interval Plot**
Prepare data for plotting
```{r}
#Rename columns for rbind
pd_1 <- rename(pd_1, "Region" ="Continent")
pd_2 <- rename(pd_2, "Region" ="L2s")
pd_3 <- rename(pd_3, "Region" ="L3s")
pd_4 <- rename(pd_4, "Region" ="L4s")

#Bind all pd objects
pd <- rbind(pd_1, pd_2)
pd <- rbind(pd, pd_3)
pd <- rbind(pd, pd_4)

#Attach p-values
#These p-values came from objects test, test2, test3, test4
pd$pv <-""
pd <- setDT(pd)[Scale == "L1", pv:="5.304e-05"]
pd <- setDT(pd)[Scale == "L2", pv:="4.279e-05"]
pd <- setDT(pd)[Scale == "L3", pv:="2.012e-08"]
pd <- setDT(pd)[Scale == "L4", pv:="0.2558"]

#Attach sample sizes
pd <- pd %>% 
  group_by(Scale) %>% mutate(n = n()) %>% 
  mutate(label = paste0(Scale))

#View confidence intervals
mean_cl_normal <- function(data, alpha = 0.05) {
  n <- length(data)
  mean_value <- mean(data)
  std_dev <- sd(data)
  stderr <- std_dev / sqrt(n)
  z <- qnorm(1 - alpha/2)
  lower <- mean_value - z * stderr
  upper <- mean_value + z * stderr
  ci <- c(lower, upper)
  result <- list(mean = mean_value, ci = ci)
  return(result)
}

mean_cl_normal(pd_1$InvasionRate)
mean_cl_normal(pd_2$InvasionRate)
mean_cl_normal(pd_3$InvasionRate)
mean_cl_normal(pd_4$InvasionRate)
```

Build and export plot (Fig. 1 Panel E)
```{r}
#Export plot
tiff("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/Figures and Figure Data/TensRule/test2.tiff",
     units = "in",
     width = 1.2671,
     height = 3.321,
     res = 800)

#build plot
g <- ggplot(pd, aes(x=`label`, y=`InvasionRate`, colour=`Scale`))+ 
                stat_summary(geom="point", fun=mean, size= 4 ) +
                stat_summary(geom="errorbar", fun.data=mean_cl_normal, width=0.7, linewidth=1.5) +
                coord_cartesian(ylim=c(0, 0.32),xlim = c(0.5,4.5), expand = FALSE, clip = "off") +
                xlab("Scale") +
                ylab("Invasion Rate (%)") +
                theme_bw()+
                theme(legend.position = "none",
                       axis.text.x = (axis.text=element_text(size=9)),
                       axis.text.y = (axis.text=element_text(size=9)), 
                       axis.title.x = element_text(face="bold", size = 10),
                       axis.title.y = element_text(face="bold", size = 10)) +
                geom_hline(yintercept= 0.1, linetype="dashed", color = "black", linewidth=1)

#View plot
g + scale_color_manual(values=c("#E7533A", "#FC8D59", "#FC8D59", "#FDC38D"))

dev.off()
```

Significance Testing
```{r}
All2 <- All2Destinations[,c(1,5)]
All2$Level <- "2"
All2 <- All2[,-c(1)]

All3 <- All3Destinations[,c(1,5)]
All3$Level <- "3"
All3 <- All3[,-c(1)]

All4 <- All4Destinations[,c(1,5)]
All4$Level <- "4"
All4 <- All4[,-c(1)]

All1 <- AllContsDestinations[,c(1,6)]
All1$Level <- "1"
All1 <- All1[,-c(1)]

#Are the proportions for each level significantly different than each other?
t.test(All1$InvasionRate, All2$InvasionRate) # ns
t.test(All1$InvasionRate, All3$InvasionRate) # **
t.test(All1$InvasionRate, All4$InvasionRate) # *** 
t.test(All2$InvasionRate, All3$InvasionRate) # ns
t.test(All2$InvasionRate, All4$InvasionRate) # **
t.test(All3$InvasionRate, All4$InvasionRate) # **

#Non-parametric way of doing the same thing: Wilcoxon test across scales
pairwise.wilcox.test(pd$InvasionRate, pd$Scale, p.adjust.method = "bonferroni")

#Both tests agree that 1 and 2 are not significant.
#t-test thinks 2-3 is not significant, wilcoxon thinks it is.
#Both tests agree the remaining combinations are significant. 


#Same thing with GLM

All2_glm <- All2Destinations[,c(2,3)]
All2_glm$Level <- "2"

All3_glm <- All3Destinations[,c(2,3)]
All3_glm$Level <- "3"

All4_glm <- All4Destinations[,c(2,3)]
All4_glm$Level <- "4"

All1_glm <- AllContsDestinations[,c(3,4)]
All1_glm$Level <- "1"

all_glm <- rbind(All1_glm,All2_glm,All3_glm,All4_glm)

test_glm5 <- glm(cbind(all_glm$InvasiveSpecies, all_glm$EstablishedSpecies) ~ all_glm$Level, family = binomial(link="logit"))
summary(test_glm5)
#Looks like it thinks all combinations are significant

#Check just 1 vs 2
glm1and2 <- rbind(All1_glm,All2_glm)

test_glm1and2 <- glm(cbind(glm1and2$InvasiveSpecies, glm1and2$EstablishedSpecies) ~ glm1and2$Level, family = binomial(link="logit"))

summary(test_glm1and2)
#Thinks 1 and 2 are different, which is different from previous 2 tests


#Check just 2 and 3
glm2and3 <- rbind(All2_glm,All3_glm)

test_glm2and3 <- glm(cbind(glm2and3$InvasiveSpecies, glm2and3$EstablishedSpecies) ~ glm2and3$Level, family = binomial(link="logit"))

summary(test_glm2and3)
#Thinks 2 and 3 are different, which agrees with wilcoxon but differs with t-test. 

#All other combinations are agreed upon by all three tests: they are all significant. 

```

Compile NA regions
```{r}
require(qpcR)
NA_regions <- qpcR:::cbind.na(L2NAs, L3NAs, L4NAs)
NA_regions <- NA_regions[,c(1,4,7)]

fwrite(NA_regions, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/NA_regions_codes.csv")
```

