---
title: "Final_Climates"
author: "Will Pfadenhauer"
date: "2023-01-03"
output: html_document
---

## 1. Prepping data. No need to run this more than once. If you're just rerunning code, start at 2A below. 

A. Load data & packages
```{r}
require(data.table)
require(dplyr)
require(tidyr)
require(PropCIs)
require(ggplot2)
require(forcats)
require(scoringutils)
require(svglite)
require(PropCIs)

#Load data
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")
```

B. Need to create a list of all L4 regions
```{r}
#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,10)]
data_i <- data_i[,c(1,14)]
data_n <- data_n[,c(1,6)]

#rename for matching
data_e <- rename(data_e, "L4" ="Established_L4")
data_i <- rename(data_i, "L4" ="Invasive_L4")
data_n <- rename(data_n, "L4" ="Native_L4")

#Combine all L4 regions
full_l4 <-rbind(data_e, data_i)
full_l4 <- rbind(full_l4, data_n)

#Remove duplicates
full_l4 <- distinct(full_l4,
                    L4)

#Sort by alphabetical order
full_l4 <- full_l4[order(full_l4$L4),]

#Remove weird rows
full_l4 <- full_l4[-c(1),]

rm(data_e)
rm(data_i)
rm(data_n)
```

C. Removing overlaps
```{r}
#Un-collapse Native and Invaded regions
overlaps <- separate_rows(data, Native_L4, sep=",")
overlaps <- separate_rows(overlaps, Invasive_L4, sep=",")

#For loop to remove matching invasive combinations
a <- c(1:636)
for (i in a) {
  
overlaps <- overlaps[!((paste(full_l4[i,1])==overlaps$Native_L4) &
                          (paste(full_l4[i,1])==overlaps$Invasive_L4)),]

 message('Running Region ', i, ' of 636')
}

#Collapse Invasive_L4
overlaps_test <- overlaps %>%
  select(1:14) %>% 
  group_by(Accepted_name, Native_L4) %>%
  mutate(Invasive_L4.x = paste(Invasive_L4, collapse = ","))

overlaps_test <- distinct(overlaps_test,
                     Accepted_name,
                     Native_L4,
                     .keep_all = TRUE)

overlaps_test <- overlaps_test[,-c(14)]
overlaps_test <- rename(overlaps_test, "Invasive_L4" ="Invasive_L4.x")
rm(overlaps)

#Expand Established_L4
overlaps_test <- separate_rows(overlaps_test, Established_L4, sep=",")

#Run another for-loop to remove matching established combinations
for (i in a) {
overlaps_test <- overlaps_test[!((paste(full_l4[i,1])==overlaps_test$Native_L4) &
                          (paste(full_l4[i,1])==overlaps_test$Established_L4)),]

 message('Running Region ', i, ' of 636')
}

#Collapse Established_L4
overlaps_test2 <- overlaps_test %>%
  select(1:14) %>% 
  group_by(Accepted_name, Native_L4) %>%
  mutate(Established_L4.x = paste(Established_L4, collapse = ","))

overlaps_test2 <- distinct(overlaps_test2,
                     Accepted_name,
                     Native_L4,
                     .keep_all = TRUE)

overlaps_test2 <- overlaps_test2[,-c(10)]
overlaps_test2 <- rename(overlaps_test2, "Established_L4" ="Established_L4.x")
overlaps_test2 <-overlaps_test2[,c(1:9,14,10:13)]

#Collapse Native_L4
overlaps_test2 <- overlaps_test2 %>%
  select(1:14) %>% 
  group_by(Accepted_name) %>%
  mutate(Native_L4.x = paste(Native_L4, collapse = ","))

overlaps_test2 <- distinct(overlaps_test2,
                     Accepted_name,
                     .keep_all = TRUE)

overlaps_test2 <- overlaps_test2[,-c(6)]
overlaps_test2 <- rename(overlaps_test2, "Native_L4" ="Native_L4.x")
overlaps_test2 <-overlaps_test2[,c(1:5,14,6:13)]

#Export results
fwrite(overlaps_test2, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/January_Climate_Overlaps_Removed_Backup.csv")

rm(overlaps_test)
rm(overlaps_test2)
rm(a)
rm(i)
rm(data)
rm(full_l4)
```

## 2. If you're just rerunning code, you can start here.

A. Join regions to matching climates
```{r}
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/January_Climate_Overlaps_Removed_Backup.csv")

clim_l4 <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L4_Better_Climate_data.csv")
clim_l3 <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L3_Better_Climate_data.csv")

#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#Join regions to l4 climates
data_e <-left_join(data_e, clim_l4, by=c("Established_L4"="Level4_cod")) 
data_i <-left_join(data_i, clim_l4, by=c("Invasive_L4"="Level4_cod")) 
data_n <-left_join(data_n, clim_l4, by=c("Native_L4"="Level4_cod"))

#Join remaining established regions to l3 climates
data_e <-left_join(data_e, clim_l3, by=c("Established_L4"="LEVEL3_COD")) 
data_e <- setDT(data_e)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_e <- setDT(data_e)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

#Join remaining invaded regions to l3 climates
data_i <-left_join(data_i, clim_l3, by=c("Invasive_L4"="LEVEL3_COD")) 
data_i <- setDT(data_i)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_i <- setDT(data_i)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

#Join remaining native regions to l3 climates
data_n <-left_join(data_n, clim_l3, by=c("Native_L4"="LEVEL3_COD")) 
data_n <- setDT(data_n)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_n <- setDT(data_n)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

rm(clim_l3)
rm(clim_l4)
```

B. If you want Islands or Mainlands Removed, do these steps. Otherwise, move on to next code chunk (C). Make sure to check lines 201-204 before running this to make sure you're generating the data you want. 
```{r}
glonaf_tdwg <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Species_Lists/GloNAF/Region_GloNAF_vanKleunenetal2018Ecology.csv")

extra_islands<- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Extra_Islands.csv")

glonaf_tdwg <- glonaf_tdwg[,-c(1:3,5,7:13)]
glonaf_tdwg <- distinct(glonaf_tdwg, 
                        tdwg4,
                        .keep_all = TRUE)
all_islands <- rbind(glonaf_tdwg, extra_islands)
all_islands <- distinct(all_islands,
                        tdwg4,
                        .keep_all = TRUE)

#Join regions to l4 regions from GloNAF
data_e <-left_join(data_e, all_islands, by=c("Established_L4"="tdwg4")) 
data_i <-left_join(data_i, all_islands, by=c("Invasive_L4"="tdwg4")) 
data_n <-left_join(data_n, all_islands, by=c("Native_L4"="tdwg4"))

#If these are set to 1, removes islands. If these are set to 0, removes mainlands.
data_e <- data_e[!data_e$island=="0",]
data_i <- data_i[!data_i$island=="0",]
data_n <- data_n[!data_n$island=="0",]

data_e <- data_e[,-c(17)]
data_i <- data_i[,-c(17)]
data_n <- data_n[,-c(17)]

rm(glonaf_tdwg)
rm(all_islands)
rm(extra_islands)

```

C. Continue here.
```{r}
#Recombine MAJORITY columns into strings (invaded)
data_i <- data_i %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(invaded_climates = paste(MAJORITY.x, collapse = ","))

data_i <- distinct(data_i,
                  Accepted_name,
                  .keep_all = TRUE)
data_i <- data_i[,-c(2)]

#Recombine MAJORITY columns into strings (established)
data_e <- data_e %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(established_climates = paste(MAJORITY.x, collapse = ","))

data_e <- distinct(data_e,
                  Accepted_name,
                  .keep_all = TRUE)
data_e <- data_e[,-c(2)]

#Recombine MAJORITY columns into strings (native)
data_n <- data_n %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(native_climates = paste(MAJORITY.x, collapse = ","))

data_n <- distinct(data_n,
                  Accepted_name,
                  .keep_all = TRUE)
data_n <- data_n[,-c(2)]

#Full climate lists
name_status <- data[,c(1,2)]
full_climates <- left_join(name_status, data_n)
full_climates <- left_join(full_climates, data_e)
full_climates <- left_join(full_climates, data_i)
rm(name_status)

#Add commas to make elements easier to search
full_climates$established_climates<- paste(",",full_climates$established_climates, sep="")
full_climates$invaded_climates<- paste(",",full_climates$invaded_climates, sep="")
full_climates$native_climates<- paste(",",full_climates$native_climates, sep="")

full_climates$established_climates<- paste(full_climates$established_climates,",", sep="")
full_climates$invaded_climates<- paste(full_climates$invaded_climates,",", sep="")
full_climates$native_climates<- paste(full_climates$native_climates,",", sep="")

#Remove NAs
full_climates$native_climates <- gsub("NA", "", full_climates$native_climates)
full_climates$established_climates <- gsub("NA", "", full_climates$established_climates)
full_climates$invaded_climates <- gsub("NA", "", full_climates$invaded_climates)

rm(data_e)
rm(data_i)
rm(data_n)
```
From here, you can either move right into the next section. Or you can skip directly to section 4A if you're only interested in Climate Flows Results (Tile Plot).

## 3. Recipient Climate Results

A. Making data table for results
```{r}
#Make a Data Table to Store Destination Results
ClimatesAsDestinations <- data.table()
ClimatesAsDestinations$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
ClimatesAsDestinations$Climate <- c("Tropical", "Arid", "Temperate", "Continental", "Polar")
ClimatesAsDestinations$EstablishedSpecies <- ""
ClimatesAsDestinations$InvasiveSpecies <- ""
```

B. Count established and invasive species arriving in each climate. Make sure to check the last three lines of code in this chunk to make sure labeling your data correctly. 
```{r}
a <- c(1:5)
for (i in a) {
  
#Need species with invaded climates = whatever Climate is in Row i
Climate_I <- full_climates[grep(paste(ClimatesAsDestinations[i,1] ), full_climates$invaded_climates),]

#Remove excess columns
Climate_I <- Climate_I[,c(1)]

#Label as invasive
Climate_I$type = "1"

#Need species with Established climates = whatever  Climate is in Row i
Climate_E <- full_climates[grep(paste(ClimatesAsDestinations[i,1] ), full_climates$established_climates),]

#Remove invasive species from our established list
Climate_E <-  Climate_E[-(grep(paste(ClimatesAsDestinations[i,1] ), Climate_E$invaded_climates)),]

#Remove excess columns
Climate_E <- Climate_E[,c(1)]

#Label as established
Climate_E$type = "0"

#Rbind Invaders to Established 
Climate <- rbind(Climate_I, Climate_E)

Climate_count <- table(Climate$type)

ClimatesAsDestinations[i,3] <- paste(Climate_count[1])
ClimatesAsDestinations[i,4] <- paste(Climate_count[2])

rm(Climate)
rm(Climate_E)
rm(Climate_I)
rm(Climate_count)
  
 message('Running Climate ', i, ' of 5')

}

#Calculate Invasion Rates
ClimatesAsDestinations$EstablishedSpecies <- as.integer(ClimatesAsDestinations$EstablishedSpecies)
ClimatesAsDestinations$InvasiveSpecies <- as.integer(ClimatesAsDestinations$InvasiveSpecies)
ClimatesAsDestinations$Total <- (ClimatesAsDestinations$EstablishedSpecies + ClimatesAsDestinations$InvasiveSpecies)
ClimatesAsDestinations$InvasionRate <- (ClimatesAsDestinations$InvasiveSpecies/ClimatesAsDestinations$Total)

#Run this - or not - depending on whether you removed islands earlier
#NoI_ClimatesAsDestinations <- ClimatesAsDestinations
NoM_ClimatesAsDestinations <- ClimatesAsDestinations

```
Return to 2A and rerun all of these chunks again for No Mainlands, then No Islands, then All Data (nothing removed).

C. Plotting Invasion Rates for Recipient Climates (Point Plot with Error Bars)
```{r}
#Assign results to correct categories
NoI_ClimatesAsDestinations$IslandData <- paste("Mainlands Only")
new_row    <- data.table(`Climate` = "All")
NOIC <- rbind(NoI_ClimatesAsDestinations, new_row, fill=TRUE) 
NOIC[6,7] <- paste("Mainlands Only")
NOIC[6,5] <- as.numeric(paste(10140))
NOIC[6,4] <- as.numeric(paste(1517))
NOIC[6,6] <- NOIC[6,4] / NOIC[6,5]
NOIC <- NOIC[c(6,1,2,3,4,5),]

NoM_ClimatesAsDestinations$IslandData <- paste("Islands Only")
NOMC <- rbind(NoM_ClimatesAsDestinations, new_row, fill=TRUE)
NOMC[6,7] <- paste("Islands Only")
NOMC[6,5] <- as.numeric(paste(6921))
NOMC[6,4] <- as.numeric(paste(1295))
NOMC[6,6] <- NOMC[6,4] / NOMC[6,5]
NOMC <- NOMC[c(6,1,2,3,4,5),]

ClimatesAsDestinations$IslandData <- paste("All Data")
CAD <- rbind(ClimatesAsDestinations, new_row, fill=TRUE)
CAD[6,7] <- paste("All Data")
CAD[6,5] <- as.numeric(paste(12126))
CAD[6,4] <- as.numeric(paste(2039))
CAD[6,6] <- CAD[6,4] / CAD[6,5]
CAD <- CAD[c(6,1,2,3,4,5),]

#Combine results from island data and no island data
pd <- rbind(CAD, NOMC, NOIC)
pd$CIMax <- ""
pd$CIMin <- ""

#Calculate Confidence Intervals
a <- c(1:18)
for (i in a) {
conf <- scoreci(as.integer(paste(pd[i,4])), as.integer(paste(pd[i,5])), conf.level = 0.95)
pd[i,8] <- conf[1]$conf.int[2]
pd[i,9] <- conf[1]$conf.int[1]
}

pd$CIMax <- as.numeric(pd$CIMax)
pd$CIMin <- as.numeric(pd$CIMin)

pd$label <- paste(c(  "All",
                      "Tropical",
                      "Arid",
                      "Temperate",
                      "Continental",
                      "Polar",
                      "All_IO",
                      "Tropical_IO",
                      "Arid_IO",
                      "Temperate_IO",
                      "Continental_IO",
                      "Polar_IO",
                      "All_MO",
                      "Tropical_MO",
                      "Arid_MO",
                      "Temperate_MO",
                      "Continental_MO",
                      "Polar_MO"))

pd$label <- factor(pd$label, levels=unique(pd$label))
```

D. Printing and exporting plot
```{r}
svglite("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/Figures and Figure Data/Climate/test2.svg",
     width = 4.76,
     height = 3.4559)

#Build Plot
g <- ggplot(pd, aes(x= fct_inorder(`label`), y=`InvasionRate`, colour=`Climate`))+ 
                stat_summary(geom="point", fun=mean, size= 2.5) +
                ylab("Invasion Rate (%)") +
                coord_cartesian(ylim=c(0, 0.35),xlim = c(0.5,18.5), expand = FALSE, clip = "off") +
                theme_bw() +
                theme(legend.position = "none",
                      plot.margin = unit(c(0.25, 0.25, 3, 0.25), "lines"),
                      axis.title.x = element_blank(),
                      axis.text.x = element_blank(),
                      axis.text.y = (axis.text=element_text(size=10)),
                      axis.title.y = element_text(face="bold", size = 10)) +
                geom_hline(yintercept= 0.1, linetype="dashed", color = "black", linewidth=1.3) +
                geom_vline(aes(xintercept = 6.5)) +
                geom_vline(aes(xintercept = 12.5)) +
                annotate(geom = "text", x = seq_len(nrow(pd)), y = -0.02 + rep(c(0,-0.015,0,-0.015,0,-0.015), 3),
                         label = pd$Climate, size = 2.5) +
                annotate(geom = "text", x = 3.5 + (6 * (0:2)), y = -0.06, label = unique(pd$IslandData),
                         size = 4, fontface = "bold")
 g + 
  geom_errorbar(aes(ymin = pd$CIMin, ymax = pd$CIMax), width = 0.4, linewidth =0.75) +
  scale_color_manual(values= c("black", "#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494",
                               "black", "#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494",
                               "black", "#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494"))
dev.off()
```

E. Significance Testing for Point Plot
```{r}
pd6 <- pd[2:6,]
pd12 <- pd[8:12,]
pd18<- pd[14:18,]

#Are the proportions for each climate significantly different than each other?
pairwise.prop.test(pd6$InvasiveSpecies, pd6$Total, conf.level=0.95, p.adjust.method = "bonferroni")
pairwise.prop.test(pd12$InvasiveSpecies, pd12$Total, conf.level=0.95, p.adjust.method = "bonferroni")
pairwise.prop.test(pd18$InvasiveSpecies, pd18$Total, conf.level=0.95, p.adjust.method = "bonferroni")

```

## 4. Climate Flows Results

A. Data Tables for Climate Flows
```{r}
#Start by making data tables for results (change these names if you're using "no islands" or "no mainlands" data)

#First one will store proportions of Established/Invasive
CFs <- data.table()
CFs$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs$From_Tropical<- ""
CFs$From_Arid<- ""
CFs$From_Temperate<- ""
CFs$From_Continental<- ""
CFs$From_Polar<- ""

#Next one will store absolute numbers of established species
CFs_E <- data.table()
CFs_E$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_E$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_E$From_Tropical<- ""
CFs_E$From_Arid<- ""
CFs_E$From_Temperate<- ""
CFs_E$From_Continental<- ""
CFs_E$From_Polar<- ""

#This one will store absolute numbers of invasive species
CFs_I <- data.table()
CFs_I$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_I$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_I$From_Tropical<- ""
CFs_I$From_Arid<- ""
CFs_I$From_Temperate<- ""
CFs_I$From_Continental<- ""
CFs_I$From_Polar<- ""

#This one will store max. values of confidence intervals
CFs_CMax <- data.table()
CFs_CMax$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_CMax$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_CMax$From_Tropical<- ""
CFs_CMax$From_Arid<- ""
CFs_CMax$From_Temperate<- ""
CFs_CMax$From_Continental<- ""
CFs_CMax$From_Polar<- ""

#Last one will store min. values of confidence intervals
CFs_CMin <- data.table()
CFs_CMin$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_CMin$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_CMin$From_Tropical<- ""
CFs_CMin$From_Arid<- ""
CFs_CMin$From_Temperate<- ""
CFs_CMin$From_Continental<- ""
CFs_CMin$From_Polar<- ""
```

B. Climate Flows (Sources to Destinations)
```{r}
a <- c(1:5)
b <- c(1:5)
for (i in a) {

#Pull out species native to ith climate  
Natives <- full_climates[grep(paste(CFs[i,1]), full_climates$native_climates),]
  
for (j in b){
#Pull out established species in ith climate (which are already native to ith climate)
e_count <- nrow(Natives[grep(paste(CFs[j,1]), Natives$established_climates),])

#Pull out invasive species in ith climate (which are already native to ith climate)
i_count <- nrow(Natives[grep(paste(CFs[j,1]), Natives$invaded_climates),])

#Do a little arithmetic
CFs[j,i+2] <- as.numeric(paste(i_count/e_count))
CFs_E[j,i+2] <- as.numeric(paste(e_count))
CFs_I[j,i+2] <- as.numeric(paste(i_count))

#Calculate Confidence Intervals
conf <- scoreci(as.integer(paste(i_count)), as.integer(paste(e_count)), conf.level = 0.95)
CFs_CMax[j,i+2] <- conf[1]$conf.int[2]
CFs_CMin[j,i+2] <- conf[1]$conf.int[1]
}}

rm(Natives)

#Run these three lines if you removed islands above
#CFs_NoI<- CFs
#CFs_E_NoI<- CFs_E
#CFs_I_NoI<- CFs_I

#Run these three lines if you removed mainlands above
#CFs_NoM<- CFs
#CFs_E_NoM<- CFs_E
#CFs_I_NoM<- CFs_I

#Leave data tables with current names if you didn't remove anything. 
#But, if you do all three versions, you have to do this one last. 
```

C. Plotting Climate Flows Results (Figure S1 - Tile Plot)
```{r}
pcfd <- CFs[,-c(1)]
pcfd$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))
pcfd$From_Tropical <- as.numeric(pcfd$From_Tropical)
pcfd$From_Arid <- as.numeric(pcfd$From_Arid)
pcfd$From_Temperate <- as.numeric(pcfd$From_Temperate)
pcfd$From_Continental <- as.numeric(pcfd$From_Continental)
pcfd$From_Polar <- as.numeric(pcfd$From_Polar)
pcfd <- pcfd %>% mutate(across(where(is.numeric), round, digits=2))
mpcfd <- melt(pcfd)

m <- ggplot(mpcfd, aes(x=Climate, y=variable)) +
        geom_tile(aes(fill=value, color=as.factor(Climate), width=0.7, height=0.7), linewidth=2) +
        scale_fill_gradientn(colours=c("white"), name = "Invasion Rate (%)") +
        geom_text(aes(label=value))+
        ylab("Source Climate") +
        xlab("Recipient Climate") +
        theme(axis.title.y = element_text(face="bold"),
              axis.title.x = element_text(face="bold"),
              legend.title = element_text(face="bold"))

#Note,this order of colors doesn't actually match the climates. I changed them manually in Illustrator.
mm <- m + scale_color_manual(values=c("#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494"), guide = 'none') +
    scale_y_discrete(labels=c('Tropical', 'Arid', 'Temperate', 'Continental', 'Polar'))

```

D. Export Plot
```{r}
#Print plot
svglite("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/Figures and Figure Data/Climate/climateflowsR.svg",
     width = 10.58333,
     height = 6.770833)
mm

dev.off()
```

F. Significance Testing for Climate Flows
```{r}
#Add absolute values to mpcfd
CFs_E <- CFs_E[,-c(1)]
CFs_I <- CFs_I[,-c(1)]

CFs_E$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))
CFs_I$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))

CFs_E$From_Tropical <- as.numeric(CFs_E$From_Tropical)
CFs_E$From_Arid <- as.numeric(CFs_E$From_Arid)
CFs_E$From_Temperate <- as.numeric(CFs_E$From_Temperate)
CFs_E$From_Continental <- as.numeric(CFs_E$From_Continental)
CFs_E$From_Polar <- as.numeric(CFs_E$From_Polar)

CFs_I$From_Tropical <- as.numeric(CFs_I$From_Tropical)
CFs_I$From_Arid <- as.numeric(CFs_I$From_Arid)
CFs_I$From_Temperate <- as.numeric(CFs_I$From_Temperate)
CFs_I$From_Continental <- as.numeric(CFs_I$From_Continental)
CFs_I$From_Polar <- as.numeric(CFs_I$From_Polar)

m_CFs_E <- melt(CFs_E)
m_CFs_I <- melt(CFs_I)

st <- left_join(mpcfd, m_CFs_E, by=c("Climate"="Climate", "variable"="variable"))
st <- rename(st, "Total_Species" ="value.y")

st <- left_join(st, m_CFs_I, by=c("Climate"="Climate", "variable"="variable"))
st<- rename(st, "Invasive_Species" ="value")

#Are the proportions for each climate significantly different than each other?
pairwise.prop.test(st$Invasive_Species, st$Total_Species, conf.level=0.95)
```
