---
title: "Final_Climates"
author: "Will Pfadenhauer"
date: "2023-01-03"
output: html_document
---

## 1. Prepping data. No need to run this more than once. If you're just rerunning code, just load the packages here and then skip to 2A below. 

A. Load data & packages
```{r}
require(data.table)
require(dplyr)
require(tidyr)
require(PropCIs)
require(ggplot2)
require(forcats)
require(scoringutils)
require(svglite)
require(PropCIs)

#Load data
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Dec_Dat_Bac_WO_NA_regions.csv")
```

B. Need to create a list of all L4 regions
```{r}
#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,10)]
data_i <- data_i[,c(1,14)]
data_n <- data_n[,c(1,6)]

#rename for matching
data_e <- rename(data_e, "L4" ="Established_L4")
data_i <- rename(data_i, "L4" ="Invasive_L4")
data_n <- rename(data_n, "L4" ="Native_L4")

#Combine all L4 regions
full_l4 <-rbind(data_e, data_i)
full_l4 <- rbind(full_l4, data_n)

#Remove duplicates
full_l4 <- distinct(full_l4,
                    L4)

#Sort by alphabetical order
full_l4 <- full_l4[order(full_l4$L4),]

#Remove weird rows
full_l4 <- full_l4[-c(1),]

rm(data_e)
rm(data_i)
rm(data_n)
```

C. Removing overlaps
```{r}
#Un-collapse Native and Invaded regions
overlaps <- separate_rows(data, Native_L4, sep=",")
overlaps <- separate_rows(overlaps, Invasive_L4, sep=",")

#For loop to remove matching invasive combinations
a <- c(1:636)
for (i in a) {
  
overlaps <- overlaps[!((paste(full_l4[i,1])==overlaps$Native_L4) &
                          (paste(full_l4[i,1])==overlaps$Invasive_L4)),]

 message('Running Region ', i, ' of 636')
}

#Collapse Invasive_L4
overlaps_test <- overlaps %>%
  select(1:14) %>% 
  group_by(Accepted_name, Native_L4) %>%
  mutate(Invasive_L4.x = paste(Invasive_L4, collapse = ","))

overlaps_test <- distinct(overlaps_test,
                     Accepted_name,
                     Native_L4,
                     .keep_all = TRUE)

overlaps_test <- overlaps_test[,-c(14)]
overlaps_test <- rename(overlaps_test, "Invasive_L4" ="Invasive_L4.x")
rm(overlaps)

#Expand Established_L4
overlaps_test <- separate_rows(overlaps_test, Established_L4, sep=",")

#Run another for-loop to remove matching established combinations
for (i in a) {
overlaps_test <- overlaps_test[!((paste(full_l4[i,1])==overlaps_test$Native_L4) &
                          (paste(full_l4[i,1])==overlaps_test$Established_L4)),]

 message('Running Region ', i, ' of 636')
}

#Collapse Established_L4
overlaps_test2 <- overlaps_test %>%
  select(1:14) %>% 
  group_by(Accepted_name, Native_L4) %>%
  mutate(Established_L4.x = paste(Established_L4, collapse = ","))

overlaps_test2 <- distinct(overlaps_test2,
                     Accepted_name,
                     Native_L4,
                     .keep_all = TRUE)

overlaps_test2 <- overlaps_test2[,-c(10)]
overlaps_test2 <- rename(overlaps_test2, "Established_L4" ="Established_L4.x")
overlaps_test2 <-overlaps_test2[,c(1:9,14,10:13)]

#Collapse Native_L4
overlaps_test2 <- overlaps_test2 %>%
  select(1:14) %>% 
  group_by(Accepted_name) %>%
  mutate(Native_L4.x = paste(Native_L4, collapse = ","))

overlaps_test2 <- distinct(overlaps_test2,
                     Accepted_name,
                     .keep_all = TRUE)

overlaps_test2 <- overlaps_test2[,-c(6)]
overlaps_test2 <- rename(overlaps_test2, "Native_L4" ="Native_L4.x")
overlaps_test2 <-overlaps_test2[,c(1:5,14,6:13)]

#Export results
fwrite(overlaps_test2, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/January_Climate_Overlaps_Removed_Backup2.csv")

rm(overlaps_test)
rm(overlaps_test2)
rm(a)
rm(i)
rm(data)
rm(full_l4)
```

## 2. If you're just rerunning code, you can start here.

A. Join regions to matching climates
```{r}
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/January_Climate_Overlaps_Removed_Backup2.csv")

clim_l4 <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L4_Better_Climate_data.csv")
clim_l3 <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L3_Better_Climate_data.csv")

#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#Join regions to l4 climates
data_e <-left_join(data_e, clim_l4, by=c("Established_L4"="Level4_cod")) 
data_i <-left_join(data_i, clim_l4, by=c("Invasive_L4"="Level4_cod")) 
data_n <-left_join(data_n, clim_l4, by=c("Native_L4"="Level4_cod"))

#Join remaining established regions to l3 climates
data_e <-left_join(data_e, clim_l3, by=c("Established_L4"="LEVEL3_COD")) 
data_e <- setDT(data_e)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_e <- setDT(data_e)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

#Join remaining invaded regions to l3 climates
data_i <-left_join(data_i, clim_l3, by=c("Invasive_L4"="LEVEL3_COD")) 
data_i <- setDT(data_i)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_i <- setDT(data_i)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

#Join remaining native regions to l3 climates
data_n <-left_join(data_n, clim_l3, by=c("Native_L4"="LEVEL3_COD")) 
data_n <- setDT(data_n)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_n <- setDT(data_n)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

rm(clim_l3)
rm(clim_l4)
```

B. If you want Islands or Mainlands Removed, do these steps. Otherwise, move on to next code chunk (C). Make sure to check lines 192-199 before running this to make sure you're generating the data you want. 
```{r}
all_islands <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Islands_WITH_oceanic_continental_classes.csv")

#Join regions to l4 regions from GloNAF
data_e <-left_join(data_e, all_islands, by=c("Established_L4"="tdwg4")) 
data_i <-left_join(data_i, all_islands, by=c("Invasive_L4"="tdwg4")) 
data_n <-left_join(data_n, all_islands, by=c("Native_L4"="tdwg4"))

#If these are set to 1, keeps islands. If these are set to 0, keeps mainlands.
#data_e <- data_e[data_e$island=="0",]
#data_i <- data_i[data_i$island=="0",]
#data_n <- data_n[data_n$island=="0",]

#If you kept islands in the previous step, you might also want to select only continental or oceanic islands
#data_e <- data_e[data_e$geo=="Oceanic",]
#data_i <- data_i[data_i$geo=="Oceanic",]
#data_n <- data_n[data_n$geo=="Oceanic",]

data_e <- data_e[,-c(17:20)]
data_i <- data_i[,-c(17:20)]
data_n <- data_n[,-c(17:20)]

rm(all_islands)

```

C. Continue here.
```{r}
#Recombine MAJORITY columns into strings (invaded)
data_i <- data_i %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(invaded_climates = paste(MAJORITY.x, collapse = ","))

data_i <- distinct(data_i,
                  Accepted_name,
                  .keep_all = TRUE)
data_i <- data_i[,-c(2)]

#Recombine MAJORITY columns into strings (established)
data_e <- data_e %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(established_climates = paste(MAJORITY.x, collapse = ","))

data_e <- distinct(data_e,
                  Accepted_name,
                  .keep_all = TRUE)
data_e <- data_e[,-c(2)]

#Recombine MAJORITY columns into strings (native)
data_n <- data_n %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(native_climates = paste(MAJORITY.x, collapse = ","))

data_n <- distinct(data_n,
                  Accepted_name,
                  .keep_all = TRUE)
data_n <- data_n[,-c(2)]

#Full climate lists
name_status <- data[,c(1,2)]
full_climates <- left_join(name_status, data_n)
full_climates <- left_join(full_climates, data_e)
full_climates <- left_join(full_climates, data_i)
rm(name_status)

#Add commas to make elements easier to search
full_climates$established_climates<- paste(",",full_climates$established_climates, sep="")
full_climates$invaded_climates<- paste(",",full_climates$invaded_climates, sep="")
full_climates$native_climates<- paste(",",full_climates$native_climates, sep="")

full_climates$established_climates<- paste(full_climates$established_climates,",", sep="")
full_climates$invaded_climates<- paste(full_climates$invaded_climates,",", sep="")
full_climates$native_climates<- paste(full_climates$native_climates,",", sep="")

#Remove NAs
full_climates$native_climates <- gsub("NA", "", full_climates$native_climates)
full_climates$established_climates <- gsub("NA", "", full_climates$established_climates)
full_climates$invaded_climates <- gsub("NA", "", full_climates$invaded_climates)

#Before we remove these data objects, make sure to enter the number of species from data_e and data_i in the correct spots in section 3C below. 
data_e <- separate_rows(data_e, established_climates, sep = ",")
data_e <- data_e[!data_e$established_climates=="NA",]
data_e <- unique(data_e$Accepted_name)
length(data_e)

data_i <- separate_rows(data_i, invaded_climates, sep = ",")
data_i <- data_i[!data_i$invaded_climates=="NA",]
data_i <- unique(data_i$Accepted_name)
length(data_i)
```
Once those length values are entered below, now we can remove these objects. 
```{r}
rm(data_e)
rm(data_i)
rm(data_n)
```
From here, you can either move right into the next section. Or you can skip directly to section 4A if you're only interested in Climate Flows Results (Tile Plot).

## 3. Recipient Climate Results

A. Making data table for results
```{r}
#Make a Data Table to Store Destination Results
ClimatesAsDestinations <- data.table()
ClimatesAsDestinations$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
ClimatesAsDestinations$Climate <- c("Tropical", "Arid", "Temperate", "Continental", "Polar")
ClimatesAsDestinations$EstablishedSpecies <- ""
ClimatesAsDestinations$InvasiveSpecies <- ""
```

B. Count established and invasive species arriving in each climate. Make sure to check the last 6 lines of code in this chunk to make sure labeling your data correctly. 
```{r}
a <- c(1:5)
for (i in a) {
  
#Need species with invaded climates = whatever Climate is in Row i
Climate_I <- full_climates[grep(paste(ClimatesAsDestinations[i,1] ), full_climates$invaded_climates),]

#Remove excess columns
Climate_I <- Climate_I[,c(1)]

#Label as invasive
Climate_I$type = "1"

#Need species with Established climates = whatever  Climate is in Row i
Climate_E <- full_climates[grep(paste(ClimatesAsDestinations[i,1] ), full_climates$established_climates),]

#Remove invasive species from our established list
Climate_E <-  Climate_E[-(grep(paste(ClimatesAsDestinations[i,1] ), Climate_E$invaded_climates)),]

#Remove excess columns
Climate_E <- Climate_E[,c(1)]

#Label as established
Climate_E$type = "0"

#Rbind Invaders to Established 
Climate <- rbind(Climate_I, Climate_E)

Climate_count <- table(Climate$type)

ClimatesAsDestinations[i,3] <- paste(Climate_count[1])
ClimatesAsDestinations[i,4] <- paste(Climate_count[2])

rm(Climate)
rm(Climate_E)
rm(Climate_I)
rm(Climate_count)
  
 message('Running Climate ', i, ' of 5')

}

#Calculate Invasion Rates
ClimatesAsDestinations$EstablishedSpecies <- as.integer(ClimatesAsDestinations$EstablishedSpecies)
ClimatesAsDestinations$InvasiveSpecies <- as.integer(ClimatesAsDestinations$InvasiveSpecies)
ClimatesAsDestinations$Total <- (ClimatesAsDestinations$EstablishedSpecies + ClimatesAsDestinations$InvasiveSpecies)
ClimatesAsDestinations$InvasionRate <- (ClimatesAsDestinations$InvasiveSpecies/ClimatesAsDestinations$Total)

#Run just one of the following lines - depending on whether you removed islands/mainlands earlier
All_ClimatesAsDestinations <- ClimatesAsDestinations
#All_Islands_ClimatesAsDestinations <- ClimatesAsDestinations
#Cont_Islands_ClimatesAsDestinations <- ClimatesAsDestinations
#Oce_Islands_ClimatesAsDestinations <- ClimatesAsDestinations
#Mainlands_ClimatesAsDestinations <- ClimatesAsDestinations

```
Return to 2A and rerun all of these chunks again for Mainlands, then 3 sets of Islands, then All Data (nothing removed).

C. Plotting Invasion Rates for Recipient Climates (Point Plot with Error Bars)
```{r}
#Assign results to correct categories
All_ClimatesAsDestinations$IslandData <- paste("All Data")
new_row    <- data.table(`Climate` = "All")
ACAD <- rbind(All_ClimatesAsDestinations, new_row, fill=TRUE)
ACAD[6,7] <- paste("All Data")
ACAD[6,5] <- as.numeric(paste(11262)) # number of rows from data_e during "all" iteration
ACAD[6,4] <- as.numeric(paste(1993)) # number of rows from data_i during "all" iteration
ACAD[6,6] <- ACAD[6,4] / ACAD[6,5]
ACAD <- ACAD[c(6,1,2,3,4,5),]

All_Islands_ClimatesAsDestinations$IslandData <- paste("Islands Only")
AICD <- rbind(All_Islands_ClimatesAsDestinations, new_row, fill=TRUE)
AICD[6,7] <- paste("Islands Only")
AICD[6,5] <- as.numeric(paste(6204)) # number of rows from data_e during "islands only" iteration
AICD[6,4] <- as.numeric(paste(1226)) # number of rows from data_i during "islands only" iteration
AICD[6,6] <- AICD[6,4] / AICD[6,5]
AICD <- AICD[c(6,1,2,3,4,5),]

Cont_Islands_ClimatesAsDestinations$IslandData <- paste("Continental Islands Only")
CICD <- rbind(Cont_Islands_ClimatesAsDestinations, new_row, fill=TRUE)
CICD[6,7] <- paste("Continental Islands Only")
CICD[6,5] <- as.numeric(paste(4305)) # number of rows from data_e during "continental islands only" iteration
CICD[6,4] <- as.numeric(paste(713)) # number of rows from data_i during "continental islands only" iteration
CICD[6,6] <- CICD[6,4] / CICD[6,5]
CICD <- CICD[c(6,1,2,3,4,5),]

Oce_Islands_ClimatesAsDestinations$IslandData <- paste("Oceanic Islands Only")
OICD <- rbind(Oce_Islands_ClimatesAsDestinations, new_row, fill=TRUE)
OICD[6,7] <- paste("Oceanic Islands Only")
OICD[6,5] <- as.numeric(paste(3999)) # number of rows from data_e during "oceanic islands only" iteration
OICD[6,4] <- as.numeric(paste(950)) # number of rows from data_i during "oceanic islands only" iteration
OICD[6,6] <- OICD[6,4] / OICD[6,5]
OICD <- OICD[c(6,1,2,3,4,5),]

Mainlands_ClimatesAsDestinations$IslandData <- paste("Mainlands Only")
MC <- rbind(Mainlands_ClimatesAsDestinations, new_row, fill=TRUE) 
MC[6,7] <- paste("Mainlands Only")
MC[6,5] <- as.numeric(paste(9448)) # number of rows from data_e during mainlands iteration
MC[6,4] <- as.numeric(paste(1511)) # number of rows from data_i during mainlands iteration
MC[6,6] <- MC[6,4] / MC[6,5]
MC <- MC[c(6,1,2,3,4,5),]

#Combine results from island data and no island data
pd <- rbind(ACAD, AICD, CICD, OICD, MC)
pd$CIMax <- ""
pd$CIMin <- ""

#Calculate Confidence Intervals
a <- c(1:30)
for (i in a) {
conf <- scoreci(as.integer(paste(pd[i,4])), as.integer(paste(pd[i,5])), conf.level = 0.95)
pd[i,8] <- conf[1]$conf.int[2]
pd[i,9] <- conf[1]$conf.int[1]
}

pd$CIMax <- as.numeric(pd$CIMax)
pd$CIMin <- as.numeric(pd$CIMin)

pd$label <- paste(c(  "All",
                      "Tropical",
                      "Arid",
                      "Temperate",
                      "Continental",
                      "Polar",
                      "All_Isl",
                      "Tropical_Isl",
                      "Arid_Isl",
                      "Temperate_Isl",
                      "Continental_Isl",
                      "Polar_Isl",
                      "All_Isl_Cont",
                      "Tropical_Isl_Cont",
                      "Arid_Isl_Cont",
                      "Temperate_Isl_Cont",
                      "Continental_Isl_Cont",
                      "Polar_Isl_Cont",
                      "All_Isl_Oce",
                      "Tropical_Isl_Oce",
                      "Arid_Isl_Oce",
                      "Temperate_Isl_Oce",
                      "Continental_Isl_Oce",
                      "Polar_Isl_Oce",
                      "All_Main",
                      "Tropical_Main",
                      "Arid_Main",
                      "Temperate_Main",
                      "Continental_Main",
                      "Polar_Main"))

pd$label <- factor(pd$label, levels=unique(pd$label))
```

D. Printing and exporting plot
```{r}
svglite("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/Figures and Figure Data/Climate/Rev_Fig_3_Panel_Main_Islands5.svg",
     width = 1.9,
     height = 3.8)

#Build Plot
g <- ggplot(pd, aes(x= fct_inorder(`label`), y=`InvasionRate`, colour=`Climate`))+ 
                stat_summary(geom="point", fun=mean, size= 2.5) +
                ylab("Invasion Rate (%)") +
                coord_cartesian(ylim=c(0, 0.40),xlim = c(0.5,6.5), expand = FALSE, clip = "off") + #change xlim here to print different "panels"
                theme_bw() +
                theme(legend.position = "none",
                      plot.margin = unit(c(0.25, 0.25, 3, 0.25), "lines"),
                      axis.title.x = element_blank(),
                      axis.text.x = element_blank(),
                      axis.text.y = (axis.text=element_text(size=10)),
                      axis.title.y = element_text(face="bold", size = 10)) +
                geom_hline(yintercept= 0.1, linetype="dashed", color = "black", linewidth=1.3) +
                geom_vline(aes(xintercept = 6.5)) +
                geom_vline(aes(xintercept = 12.5)) +
                annotate(geom = "text", x = seq_len(nrow(pd)), y = -0.02 + rep(c(0,-0.015,0,-0.015,0,-0.015), 5),
                         label = pd$Climate, size = 2.5) +
                annotate(geom = "text", x = 3.5 + (6 * (0:4)), y = -0.06, label = unique(pd$IslandData),
                         size = 4, fontface = "bold")
 g + 
  geom_errorbar(aes(ymin = pd$CIMin, ymax = pd$CIMax), width = 0.4, linewidth =0.75) +
  scale_color_manual(values= c("black", "#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494",
                               "black", "#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494",
                               "black", "#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494"))
dev.off()
```

E. Significance Testing for Point Plot
```{r}
pd6 <- pd[2:6,]
pd12 <- pd[8:12,]
pd18<- pd[14:18,]
pd24<- pd[20:24,]
pd30<- pd[26:30,]

#Are the proportions for each climate significantly different than each other?
pairwise.prop.test(pd6$InvasiveSpecies, pd6$Total, conf.level=0.95, p.adjust.method = "bonferroni")
pairwise.prop.test(pd12$InvasiveSpecies, pd12$Total, conf.level=0.95, p.adjust.method = "bonferroni")
pairwise.prop.test(pd18$InvasiveSpecies, pd18$Total, conf.level=0.95, p.adjust.method = "bonferroni")
pairwise.prop.test(pd24$InvasiveSpecies, pd24$Total, conf.level=0.95, p.adjust.method = "bonferroni")
pairwise.prop.test(pd30$InvasiveSpecies, pd30$Total, conf.level=0.95, p.adjust.method = "bonferroni")
```

## 4. Climate Flows Results

A. Data Tables for Climate Flows
```{r}
#Start by making data tables for results (change these names if you're using "no islands" or "no mainlands" data)

#First one will store proportions of Established/Invasive
CFs <- data.table()
CFs$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs$From_Tropical<- ""
CFs$From_Arid<- ""
CFs$From_Temperate<- ""
CFs$From_Continental<- ""
CFs$From_Polar<- ""

#Next one will store absolute numbers of established species
CFs_E <- data.table()
CFs_E$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_E$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_E$From_Tropical<- ""
CFs_E$From_Arid<- ""
CFs_E$From_Temperate<- ""
CFs_E$From_Continental<- ""
CFs_E$From_Polar<- ""

#This one will store absolute numbers of invasive species
CFs_I <- data.table()
CFs_I$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_I$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_I$From_Tropical<- ""
CFs_I$From_Arid<- ""
CFs_I$From_Temperate<- ""
CFs_I$From_Continental<- ""
CFs_I$From_Polar<- ""

#This one will store max. values of confidence intervals
CFs_CMax <- data.table()
CFs_CMax$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_CMax$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_CMax$From_Tropical<- ""
CFs_CMax$From_Arid<- ""
CFs_CMax$From_Temperate<- ""
CFs_CMax$From_Continental<- ""
CFs_CMax$From_Polar<- ""

#Last one will store min. values of confidence intervals
CFs_CMin <- data.table()
CFs_CMin$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_CMin$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_CMin$From_Tropical<- ""
CFs_CMin$From_Arid<- ""
CFs_CMin$From_Temperate<- ""
CFs_CMin$From_Continental<- ""
CFs_CMin$From_Polar<- ""
```

B. Climate Flows (Sources to Destinations)
```{r}
a <- c(1:5)
b <- c(1:5)
for (i in a) {

#Pull out species native to ith climate  
Natives <- full_climates[grep(paste(CFs[i,1]), full_climates$native_climates),]
  
for (j in b){
#Pull out established species in ith climate (which are already native to ith climate)
e_count <- nrow(Natives[grep(paste(CFs[j,1]), Natives$established_climates),])

#Pull out invasive species in ith climate (which are already native to ith climate)
i_count <- nrow(Natives[grep(paste(CFs[j,1]), Natives$invaded_climates),])

#Do a little arithmetic
CFs[j,i+2] <- as.numeric(paste(i_count/e_count))
CFs_E[j,i+2] <- as.numeric(paste(e_count))
CFs_I[j,i+2] <- as.numeric(paste(i_count))

#Calculate Confidence Intervals
conf <- scoreci(as.integer(paste(i_count)), as.integer(paste(e_count)), conf.level = 0.95)
CFs_CMax[j,i+2] <- conf[1]$conf.int[2]
CFs_CMin[j,i+2] <- conf[1]$conf.int[1]
}}

rm(Natives)

#Run these three lines if you removed islands above
#CFs_NoI<- CFs
#CFs_E_NoI<- CFs_E
#CFs_I_NoI<- CFs_I

#Run these three lines if you removed mainlands above
#CFs_NoM<- CFs
#CFs_E_NoM<- CFs_E
#CFs_I_NoM<- CFs_I

#Leave data tables with current names if you didn't remove anything. 
#But, if you do all three versions, you have to do this one last. 
```

C. Plotting Climate Flows Results (Figure S1 - Tile Plot)
```{r}
pcfd <- CFs[,-c(1)]
pcfd$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))
pcfd$From_Tropical <- as.numeric(pcfd$From_Tropical)
pcfd$From_Arid <- as.numeric(pcfd$From_Arid)
pcfd$From_Temperate <- as.numeric(pcfd$From_Temperate)
pcfd$From_Continental <- as.numeric(pcfd$From_Continental)
pcfd$From_Polar <- as.numeric(pcfd$From_Polar)
pcfd <- pcfd %>% mutate(across(where(is.numeric), round, digits=2))
mpcfd <- melt(pcfd)

m <- ggplot(mpcfd, aes(x=Climate, y=variable)) +
        geom_tile(aes(fill=value, color=as.factor(Climate), width=0.7, height=0.7), linewidth=2) +
        scale_fill_gradientn(colours=c("white"), name = "Invasion Rate (%)") +
        geom_text(aes(label=value))+
        ylab("Source Climate") +
        xlab("Recipient Climate") +
        theme(axis.title.y = element_text(face="bold"),
              axis.title.x = element_text(face="bold"),
              legend.title = element_text(face="bold"))

#Note,this order of colors doesn't actually match the climates. I changed them manually in Illustrator.
mm <- m + scale_color_manual(values=c("#2c7fb8", "#a1dab4", "#ffffcc", "#41b6c4", "#253494"), guide = 'none') +
    scale_y_discrete(labels=c('Tropical', 'Arid', 'Temperate', 'Continental', 'Polar'))

```

D. Export Plot
```{r}
#Print plot
svglite("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/Figures and Figure Data/Climate/climateflowsR2.svg",
     width = 10.58333,
     height = 6.770833)
mm

dev.off()
```

F. Significance Testing for Climate Flows
```{r}
#Add absolute values to mpcfd
CFs_E <- CFs_E[,-c(1)]
CFs_I <- CFs_I[,-c(1)]

CFs_E$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))
CFs_I$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))

CFs_E$From_Tropical <- as.numeric(CFs_E$From_Tropical)
CFs_E$From_Arid <- as.numeric(CFs_E$From_Arid)
CFs_E$From_Temperate <- as.numeric(CFs_E$From_Temperate)
CFs_E$From_Continental <- as.numeric(CFs_E$From_Continental)
CFs_E$From_Polar <- as.numeric(CFs_E$From_Polar)

CFs_I$From_Tropical <- as.numeric(CFs_I$From_Tropical)
CFs_I$From_Arid <- as.numeric(CFs_I$From_Arid)
CFs_I$From_Temperate <- as.numeric(CFs_I$From_Temperate)
CFs_I$From_Continental <- as.numeric(CFs_I$From_Continental)
CFs_I$From_Polar <- as.numeric(CFs_I$From_Polar)

m_CFs_E <- melt(CFs_E)
m_CFs_I <- melt(CFs_I)

st <- left_join(mpcfd, m_CFs_E, by=c("Climate"="Climate", "variable"="variable"))
st <- rename(st, "Total_Species" ="value.y")

st <- left_join(st, m_CFs_I, by=c("Climate"="Climate", "variable"="variable"))
st<- rename(st, "Invasive_Species" ="value")

#Are the proportions for each climate significantly different than each other?
pairwise.prop.test(st$Invasive_Species, st$Total_Species, conf.level=0.95)
```
