---
title: "Final_Islands"
author: "Will Pfadenhauer"
date: "2023-01-11"
output: html_document
---

Load data & packages
```{r}
require(data.table)
require(dplyr)
require(tidyr)
require(ggplot2)
require(PropCIs)
require(forcats)
require(scoringutils)
require(Hmisc)
require(svglite)
require(naniar)
require(weights)

#Load data
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Dec_Dat_Bac_WO_NA_regions.csv")
```

## Islands vs. Mainlands at Global Scale

```{r}
#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
```

No need to re-run this part, just load the islands file at the beginning of the next chunk. 
This just shows how I got the file that I edited manually with Oceanic/Continental statuses. 
```{r}
#glonaf_tdwg <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Species_Lists/GloNAF/Region_GloNAF_vanKleunenetal2018Ecology.csv")

#extra_islands<- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Extra_Islands.csv")

#glonaf_tdwg <- glonaf_tdwg[,-c(1:3,5,7:13)]
#glonaf_tdwg <- distinct(glonaf_tdwg, 
#                        tdwg4,
#                        .keep_all = TRUE)
#all_islands <- rbind(glonaf_tdwg, extra_islands)
#all_islands <- distinct(all_islands,
#                        tdwg4,
#                        .keep_all = TRUE)

#Export for manual classification of oceanic/continental islands
#fwrite(all_islands, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Islands_needing_oceanic_continental_classes.csv")
```


```{r}
#Load in completed islands list
all_islands <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Islands_WITH_oceanic_continental_classes.csv")

#Join regions to l4 regions from GloNAF
data_e <-left_join(data_e, all_islands, by=c("Established_L4"="tdwg4")) 
data_i <-left_join(data_i, all_islands, by=c("Invasive_L4"="tdwg4")) 

#Remove mainlands.
data_e_i <- data_e[!data_e$island=="0",]
data_i_i <- data_i[!data_i$island=="0",]

data_e_i_o <- data_e_i[!data_e_i$geo=="Continental",]  #data_established_islands_oceanic
data_e_i_c <- data_e_i[!data_e_i$geo=="Oceanic",]  #data_established_islands_continental
  
data_i_i_o <- data_i_i[!data_i_i$geo=="Continental",]  #data_invasive_islands_oceanic
data_i_i_c <- data_i_i[!data_i_i$geo=="Oceanic",]  #data_invasive_islands_continental

#Remove islands.
data_e_m <- data_e[!data_e$island=="1",]
data_i_m <- data_i[!data_i$island=="1",]

#Remove duplicate species rows
data_e_i <- distinct(data_e_i, Accepted_name, .keep_all = TRUE)
data_i_i <- distinct(data_i_i, Accepted_name, .keep_all = TRUE)
data_e_m <- distinct(data_e_m, Accepted_name, .keep_all = TRUE)
data_i_m <- distinct(data_i_m, Accepted_name, .keep_all = TRUE) 

data_e_i_o <- distinct(data_e_i_o, Accepted_name, .keep_all = TRUE) 
data_e_i_c <- distinct(data_e_i_c, Accepted_name, .keep_all = TRUE) 
data_i_i_o <- distinct(data_i_i_o, Accepted_name, .keep_all = TRUE) 
data_i_i_c <- distinct(data_i_i_c, Accepted_name, .keep_all = TRUE) 


rm(data_e)
rm(data_i)
rm(glonaf_tdwg)
rm(all_islands)
rm(extra_islands)
```

```{r}
Global <- data.table(group = c("All Islands","Oceanic Islands","Continental Islands","Mainlands"), TotalSpecies = c("","","",""), InvasiveSpecies= c("","","",""))

#Transfer results to data table
Global[1,2] <- paste(nrow(data_e_i))
Global[1,3] <- paste(nrow(data_i_i))
Global[2,2] <- paste(nrow(data_e_i_o))
Global[2,3] <- paste(nrow(data_i_i_o))
Global[3,2] <- paste(nrow(data_e_i_c))
Global[3,3] <- paste(nrow(data_i_i_c))
Global[4,2] <- paste(nrow(data_e_m))
Global[4,3] <- paste(nrow(data_i_m))  

#Calculate Invasion Rate
Global$InvasiveSpecies <- as.numeric(Global$InvasiveSpecies)
Global$TotalSpecies <- as.numeric(Global$TotalSpecies)
Global$InvasionRate <- (Global$InvasiveSpecies/Global$TotalSpecies)

#Calculate Confidence intervals
Global$CIMax <- ""
Global$CIMin <- ""

a <- c(1:4)
for (i in a) {
conf <- scoreci(as.integer(paste(Global[i,3])), as.integer(paste(Global[i,2])), conf.level = 0.95)
Global[i,5] <- conf[1]$conf.int[2]
Global[i,6] <- conf[1]$conf.int[1]
}

Global$CIMax <- as.numeric(Global$CIMax)
Global$CIMin <- as.numeric(Global$CIMin)

Global$scale <- "global"
```

## Islands vs. Mainlands at WGSRPD Level 3 Scale (Countries/States)

```{r}
#Need to create a list of all L3 regions
#Expand L3 regions
data_e <- separate_rows(data, Established_L3, sep = ",")
data_i <- separate_rows(data, Invasive_L3, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,9)]
data_i <- data_i[,c(1,13)]

#rename for matching
data_e <- rename(data_e, "L3" ="Established_L3")
data_i <- rename(data_i, "L3" ="Invasive_L3")

#Combine all L3 regions
full_l3 <-rbind(data_e, data_i)

#Remove duplicates
full_l3 <- distinct(full_l3,
                    L3)

#Sort by alphabetical order
full_l3 <- full_l3[order(full_l3$L3),]

#Remove weird rows
full_l3 <- full_l3[-c(1),]
```

```{r}
#Make a Data Table to Store L3 Results
All3Destinations <- data.table()
All3Destinations$L3s <- paste(",", full_l3$L3, ",", sep = "")
All3Destinations$EstablishedSpecies <- ""
All3Destinations$InvasiveSpecies <- ""
```


```{r}
#Count established and invasive species for each L3 region
b <- c(1:251)
for (i in b) {
  
  #Need species with Invasive_L3 = whatever L3 region is in Row i
  Region_I <- data[grep(paste(All3Destinations[i,1] ), data$Invasive_L3),]
  
  #Remove species with Native L3 = whatever L3 region is in Row i 
  Region_I <- Region_I[-(grep(paste(All3Destinations[i,1]), Region_I$Native_L3)), ]
  
  #Remove excess columns
  Region_I <- Region_I[,c(1)]
  
  #Label as invasive
  Region_I$type = "1"
  
  #Need species with Established_L3 = whatever L3 region is in Row i
  Region_E <- data[grep(paste(All3Destinations[i,1] ), data$Established_L3),]
  
  #Remove species with Native L3 = whatever L3 region is in Row i
  Region_E <- Region_E[-(grep(paste(All3Destinations[i,1]), Region_E$Native_L3)),]
  
  #Remove invasive species from our established list
  Region_E <-  Region_E[-(grep(paste(All3Destinations[i,1] ), Region_E$Invasive_L3)), ]
  
  #Remove excess columns
  Region_E <- Region_E[,c(1)]
  
  #Label as established
  Region_E$type = "0"
  
  #Rbind Invaders to Established 
  Region <- rbind(Region_I, Region_E)
  
  Region_count <- table(Region$type)
  
  All3Destinations[i,2] <- paste(Region_count[1])
  All3Destinations[i,3] <- paste(Region_count[2])
  
  message('Running Region ', i, ' of 251')
  
  rm(Region)
  rm(Region_E)
  rm(Region_I)
  rm(Region_count)
  
}
```

```{r}
#Remove NAs
All3Destinations <- as.data.frame(All3Destinations)
All3Destinations <- replace_with_na(All3Destinations, replace = list(EstablishedSpecies = c("NA")))
All3Destinations <- replace_with_na(All3Destinations, replace = list(InvasiveSpecies = c("NA")))
All3Destinations <- na.omit(All3Destinations)

#Calculate Invasion Rates
All3Destinations$EstablishedSpecies <- as.integer(All3Destinations$EstablishedSpecies)
All3Destinations$InvasiveSpecies <- as.integer(All3Destinations$InvasiveSpecies)
All3Destinations$Total <- (All3Destinations$EstablishedSpecies + All3Destinations$InvasiveSpecies)

#Only use this line of code if you're doing the version of the analysis presented in Table S5. 
All3Destinations <- subset(All3Destinations, Total >= 50)

All3Destinations$InvasionRate <- (All3Destinations$InvasiveSpecies/All3Destinations$Total)
All3Destinations$Weight <- All3Destinations$Total/12352

#Prepare data for plotting
pd_3 <- All3Destinations[,c(1,5,6)]
pd_3$Scale <- "L3"

#Load Island data
all_islands <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Islands_WITH_oceanic_continental_classes.csv")

#Only retain L3 regions (remove L4 ones)
all_islands_L3 <- all_islands[grep(paste("-OO"), all_islands$tdwg4),]

#Remove extra characters so it will match pd_3
all_islands_L3$tdwg4 <- gsub("-OO", "", all_islands_L3$tdwg4)

#Add commas so it will match pd_3
all_islands_L3$tdwg4 <- paste(",", all_islands_L3$tdwg4, ",", sep = "")

#Join L3 invasion rates to island data
pd_3 <-left_join(pd_3, all_islands_L3, by=c("L3s"="tdwg4")) 

#Separate Islands and Mainlands
pd_3_i <- pd_3[!pd_3$island =="0",]
pd_3_i_o <- pd_3_i[!pd_3_i$geo =="Continental",]
pd_3_i_c <- pd_3_i[!pd_3_i$geo =="Oceanic",]
pd_3_m <- pd_3[!pd_3$island =="1",]  
```

Use this code chunk for main text results (if you use this chunk, skip the next one!)
```{r}
#Weighted average for islands (this version used in main text)
isl_w_mean<- (sum((pd_3_i$InvasionRate * pd_3_i$Weight)))/sum(pd_3_i$Weight)
isl_oce_w_mean<- (sum((pd_3_i_o$InvasionRate * pd_3_i_o$Weight)))/sum(pd_3_i_o$Weight)
isl_con_w_mean<- (sum((pd_3_i_c$InvasionRate * pd_3_i_c$Weight)))/sum(pd_3_i_c$Weight)
mld_w_mean <- (sum((pd_3_m$InvasionRate * pd_3_m$Weight)))/sum(pd_3_m$Weight)

#Calculate confidence intervals for weighted averages
weighted.ttest.ci <- function(x, weights, conf.level = 0.95) {
    nx <- length(x)
    df <- nx - 1
    vx <- wtd.var(x, weights, normwt = TRUE) 
    mx <- weighted.mean(x, weights)
    stderr <- sqrt(vx/nx)
    tstat <- mx/stderr 
    alpha <- 1 - conf.level
    cint <- qt(1 - alpha/2, df)
    cint <- tstat + c(-cint, cint)
    cint * stderr
}

#Calculate confidence intervals
isl_w_conf <- weighted.ttest.ci(pd_3_i$InvasionRate, pd_3_i$Weight)
isl_oce_w_conf <- weighted.ttest.ci(pd_3_i_o$InvasionRate, pd_3_i_o$Weight)
isl_con_w_conf <- weighted.ttest.ci(pd_3_i_c$InvasionRate, pd_3_i_c$Weight)
mld_w_conf <- weighted.ttest.ci(pd_3_m$InvasionRate, pd_3_m$Weight)
```

Use this code chunk for Table S5 (alternative version of results; if you use this chunk, skip the previous one!)
```{r}
#Averages for islands (this version used in Table S5)
isl_w_mean<- mean(pd_3_i$InvasionRate)
isl_oce_w_mean<- mean(pd_3_i_o$InvasionRate)
isl_con_w_mean<- mean(pd_3_i_c$InvasionRate)
mld_w_mean <- mean(pd_3_m$InvasionRate)

#View confidence intervals
mean_cl_normal <- function(data, alpha = 0.05) {
  n <- length(data)
  mean_value <- mean(data)
  std_dev <- sd(data)
  stderr <- std_dev / sqrt(n)
  z <- qnorm(1 - alpha/2)
  lower <- mean_value - z * stderr
  upper <- mean_value + z * stderr
  ci <- c(lower, upper)
  result <- list(mean = mean_value, ci = ci)
  return(result)
}

#Calculate confidence intervals
isl_w_conf <- mean_cl_normal(pd_3_i$InvasionRate)
isl_oce_w_conf <- mean_cl_normal(pd_3_i_o$InvasionRate)
isl_con_w_conf <- mean_cl_normal(pd_3_i_c$InvasionRate)
mld_w_conf <- mean_cl_normal(pd_3_m$InvasionRate)
```


```{r}
#Island plot with both L3 and Global scales
#Prepare data for plotting
Global2<-Global
Global2[,c(2:7)]<-""
Global <- rbind(Global,Global2)

#Fill with new data
Global[5,4] <- isl_w_mean
Global[5,5] <- isl_w_conf[2]
Global[5,6] <- isl_w_conf[1]
Global[5,7] <- "L3"

Global[6,4] <- isl_oce_w_mean
Global[6,5] <- isl_oce_w_conf[2]
Global[6,6] <- isl_oce_w_conf[1]
Global[6,7] <- "L3"

Global[7,4] <- isl_con_w_mean
Global[7,5] <- isl_con_w_conf[2]
Global[7,6] <- isl_con_w_conf[1]
Global[7,7] <- "L3"

Global[8,4] <- mld_w_mean
Global[8,5] <- mld_w_conf[2]
Global[8,6] <- mld_w_conf[1]
Global[8,7] <- "L3"


#Create labels
Global$label <- paste(c("Global All Islands",
                        "Global Oceanic Islands",
                        "Global Continental Islands",
                        "Global Mainlands",
                        "L3 All Islands",
                        "L3 Oceanic Islands",
                        "L3 Continental Islands",
                        "L3 Mainlands"))

Global$label <- factor(Global$label, levels=unique(Global$label))
Global$colorgroups <- paste(c("colorone",
                        "colortwo",
                        "colortwo",
                        "colorthree",
                        "colorone",
                        "colortwo",
                        "colortwo",
                        "colorthree"))
```

```{r}
#Build Plot
g <- ggplot(Global, aes(x= fct_inorder(`label`), y=`InvasionRate`, colour=`colorgroups`))+ 
  stat_summary(geom="point", fun=mean, size= 5 ) +
  ylab("Invasion Rate (%)") +
  coord_cartesian(ylim=c(0, 0.31),xlim = c(0.5, 8.5), expand = FALSE, clip = "off") +
  theme_bw() +
  theme(legend.position = "none",
        plot.margin = unit(c(1, 1, 3, 1), "lines"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_text(face="bold", size = 14),
        axis.text.y = element_text(size = 16)) +
  geom_hline(yintercept= 0.1, linetype="dashed", color = "black", linewidth=2) +
  geom_vline(aes(xintercept = 4.5)) +
  annotate(geom = "text", x = seq_len(nrow(Global)), y = -0.01,
           label = Global$group, size = 2.5) +
  annotate(geom = "text", x = 2.5 + (4 * (0:1)), y = -0.025, label = unique(Global$scale),
           size = 5, fontface = "bold")
p <- g + 
  geom_errorbar(aes(ymin = Global$CIMin, ymax = Global$CIMax), width = 0.3, linewidth =1.5) 
```

```{r}
#Print plot
svglite("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/Figures and Figure Data/Islands/islandwtd_OC.svg",
     width = 6,
     height = 6.072917)

p

dev.off()
```


Significance Testing
```{r}
#Global scale comparison
#Need to do a one-tailed test since the hypothesis is that islands are MORE susceptible than mainlands. 

#All islands vs. Mainlands
GST <- Global[c(1,4),]
prop.test(GST$InvasiveSpecies, GST$TotalSpecies, alternative = c("greater"), conf.level=0.95)
#significant

#Oceanic Islands vs. Mainlands
GST <- Global[c(2,4),]
prop.test(GST$InvasiveSpecies, GST$TotalSpecies, alternative = c("greater"), conf.level=0.95)
#significant

#Cont. Islands vs. Mainlands
GST <- Global[c(3,4),]
prop.test(GST$InvasiveSpecies, GST$TotalSpecies, alternative = c("greater"), conf.level=0.95)
# NOT significant


#WGSRPD Level 3 scale comparison
#All islands vs. mainlands
L3STI <- pd_3[pd_3$island =="1",]
L3STM <- pd_3[pd_3$island =="0",]

wtd.t.test(L3STI$InvasionRate,
           L3STM$InvasionRate, 
           weight = L3STI$Weight, 
           weighty = L3STM$Weight,
           samedata = FALSE,
           alternative = "greater")
#significant

#Oceanic islands vs. mainlands
L3STI_O <- L3STI[L3STI$geo =="Oceanic",]

wtd.t.test(L3STI_O$InvasionRate,
           L3STM$InvasionRate, 
           weight = L3STI_O$Weight, 
           weighty = L3STM$Weight,
           samedata = FALSE,
           alternative = "greater")
#significant

#Cont. islands vs. mainlands
L3STI_C <- L3STI[L3STI$geo =="Continental",]

wtd.t.test(L3STI_C$InvasionRate,
           L3STM$InvasionRate, 
           weight = L3STI_C$Weight, 
           weighty = L3STM$Weight,
           samedata = FALSE,
           alternative = "greater")
#NOT significant
```

Second version of L3 significance testing (this version for Table S5)
```{r}
#WGSRPD Level 3 scale comparison
#All islands vs. mainlands
L3STI <- pd_3[pd_3$island =="1",]
L3STM <- pd_3[pd_3$island =="0",]

t.test(L3STI$InvasionRate, L3STM$InvasionRate, alternative = "greater")
#significant

#Oceanic islands vs. mainlands
L3STI_O <- L3STI[L3STI$geo =="Oceanic",]

t.test(L3STI_O$InvasionRate, L3STM$InvasionRate, alternative = "greater")
#significant

#Cont. islands vs. mainlands
L3STI_C <- L3STI[L3STI$geo =="Continental",]
t.test(L3STI_C$InvasionRate, L3STM$InvasionRate, alternative = "greater")

```

