---
title: "Final_Climates"
author: "Will Pfadenhauer"
date: "2023-01-03"
output: html_document
---

## 1. Prepping data. No need to run this more than once.

Load data & packages

```{r}
require(data.table)
require(dplyr)
require(tidyr)
require(PropCIs)
require(ggplot2)
require(forcats)
require(scoringutils)

#Load data
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")
```

Need to create a list of all L4 regions

```{r}
#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#remove extra columns
data_e <- data_e[,c(1,10)]
data_i <- data_i[,c(1,14)]
data_n <- data_n[,c(1,6)]

#rename for matching
data_e <- rename(data_e, "L4" ="Established_L4")
data_i <- rename(data_i, "L4" ="Invasive_L4")
data_n <- rename(data_n, "L4" ="Native_L4")

#Combine all L4 regions
full_l4 <-rbind(data_e, data_i)
full_l4 <- rbind(full_l4, data_n)

#Remove duplicates
full_l4 <- distinct(full_l4,
                    L4)

#Sort by alphabetical order
full_l4 <- full_l4[order(full_l4$L4),]

#Remove weird rows
full_l4 <- full_l4[-c(1),]

rm(data_e)
rm(data_i)
rm(data_n)
```

Removing overlaps

```{r}
#Un-collapse Native and Invaded regions
overlaps <- separate_rows(data, Native_L4, sep=",")
overlaps <- separate_rows(overlaps, Invasive_L4, sep=",")

#For loop to remove matching invasive combinations
a <- c(1:636)
for (i in a) {
  
overlaps <- overlaps[!((paste(full_l4[i,1])==overlaps$Native_L4) &
                          (paste(full_l4[i,1])==overlaps$Invasive_L4)),]

 message('Running Region ', i, ' of 636')
}

#Collapse Invasive_L4
overlaps_test <- overlaps %>%
  select(1:14) %>% 
  group_by(Accepted_name, Native_L4) %>%
  mutate(Invasive_L4.x = paste(Invasive_L4, collapse = ","))

overlaps_test <- distinct(overlaps_test,
                     Accepted_name,
                     Native_L4,
                     .keep_all = TRUE)

overlaps_test <- overlaps_test[,-c(14)]
overlaps_test <- rename(overlaps_test, "Invasive_L4" ="Invasive_L4.x")
rm(overlaps)

#Expand Established_L4
overlaps_test <- separate_rows(overlaps_test, Established_L4, sep=",")

#Run another for-loop to remove matching established combinations
for (i in a) {
overlaps_test <- overlaps_test[!((paste(full_l4[i,1])==overlaps_test$Native_L4) &
                          (paste(full_l4[i,1])==overlaps_test$Established_L4)),]

 message('Running Region ', i, ' of 636')
}

#Collapse Established_L4
overlaps_test2 <- overlaps_test %>%
  select(1:14) %>% 
  group_by(Accepted_name, Native_L4) %>%
  mutate(Established_L4.x = paste(Established_L4, collapse = ","))

overlaps_test2 <- distinct(overlaps_test2,
                     Accepted_name,
                     Native_L4,
                     .keep_all = TRUE)

overlaps_test2 <- overlaps_test2[,-c(10)]
overlaps_test2 <- rename(overlaps_test2, "Established_L4" ="Established_L4.x")
overlaps_test2 <-overlaps_test2[,c(1:9,14,10:13)]

#Collapse Native_L4
overlaps_test2 <- overlaps_test2 %>%
  select(1:14) %>% 
  group_by(Accepted_name) %>%
  mutate(Native_L4.x = paste(Native_L4, collapse = ","))

overlaps_test2 <- distinct(overlaps_test2,
                     Accepted_name,
                     .keep_all = TRUE)

overlaps_test2 <- overlaps_test2[,-c(6)]
overlaps_test2 <- rename(overlaps_test2, "Native_L4" ="Native_L4.x")
overlaps_test2 <-overlaps_test2[,c(1:5,14,6:13)]

#Export results
fwrite(overlaps_test2, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/January_Climate_Overlaps_Removed_Backup.csv")

rm(overlaps_test)
rm(overlaps_test2)
rm(a)
rm(i)
rm(data)
rm(full_l4)
```

## 2. If you're just rerunning code, you can start here.

Join regions to matching climates

```{r}
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/January_Climate_Overlaps_Removed_Backup.csv")

clim_l4 <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L4_Better_Climate_data.csv")
clim_l3 <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L3_Better_Climate_data.csv")

#Expand L4 regions
data_e <- separate_rows(data, Established_L4, sep = ",")
data_i <- separate_rows(data, Invasive_L4, sep = ",")
data_n <- separate_rows(data, Native_L4, sep = ",")

#Join regions to l4 climates
data_e <-left_join(data_e, clim_l4, by=c("Established_L4"="Level4_cod")) 
data_i <-left_join(data_i, clim_l4, by=c("Invasive_L4"="Level4_cod")) 
data_n <-left_join(data_n, clim_l4, by=c("Native_L4"="Level4_cod"))

#Join remaining established regions to l3 climates
data_e <-left_join(data_e, clim_l3, by=c("Established_L4"="LEVEL3_COD")) 
data_e <- setDT(data_e)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_e <- setDT(data_e)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

#Join remaining invaded regions to l3 climates
data_i <-left_join(data_i, clim_l3, by=c("Invasive_L4"="LEVEL3_COD")) 
data_i <- setDT(data_i)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_i <- setDT(data_i)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

#Join remaining native regions to l3 climates
data_n <-left_join(data_n, clim_l3, by=c("Native_L4"="LEVEL3_COD")) 
data_n <- setDT(data_n)[MAJORITY.x == "", MAJORITY.x:= MAJORITY.y]
data_n <- setDT(data_n)[is.na(MAJORITY.x), MAJORITY.x:= MAJORITY.y]

rm(clim_l3)
rm(clim_l4)
```

If you want Islands or Mainlands Removed, do these steps. Otherwise, move on to next code chunk.

```{r}
glonaf_tdwg <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Species_Lists/GloNAF/Region_GloNAF_vanKleunenetal2018Ecology.csv")

extra_islands<- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Extra_Islands.csv")

glonaf_tdwg <- glonaf_tdwg[,-c(1:3,5,7:13)]
glonaf_tdwg <- distinct(glonaf_tdwg, 
                        tdwg4,
                        .keep_all = TRUE)
all_islands <- rbind(glonaf_tdwg, extra_islands)
all_islands <- distinct(all_islands,
                        tdwg4,
                        .keep_all = TRUE)

#Join regions to l4 regions from GloNAF
data_e <-left_join(data_e, all_islands, by=c("Established_L4"="tdwg4")) 
data_i <-left_join(data_i, all_islands, by=c("Invasive_L4"="tdwg4")) 
data_n <-left_join(data_n, all_islands, by=c("Native_L4"="tdwg4"))

#If these are set to 1, removes islands. If these are set to 0, removes mainlands.
data_e <- data_e[!data_e$island=="0",]
data_i <- data_i[!data_i$island=="0",]
data_n <- data_n[!data_n$island=="0",]

data_e <- data_e[,-c(17)]
data_i <- data_i[,-c(17)]
data_n <- data_n[,-c(17)]

rm(glonaf_tdwg)
rm(all_islands)
rm(extra_islands)

```

Continue here.

```{r}
#Recombine MAJORITY columns into strings (invaded)
data_i <- data_i %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(invaded_climates = paste(MAJORITY.x, collapse = ","))

data_i <- distinct(data_i,
                  Accepted_name,
                  .keep_all = TRUE)
data_i <- data_i[,-c(2)]

#Recombine MAJORITY columns into strings (established)
data_e <- data_e %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(established_climates = paste(MAJORITY.x, collapse = ","))

data_e <- distinct(data_e,
                  Accepted_name,
                  .keep_all = TRUE)
data_e <- data_e[,-c(2)]

#Recombine MAJORITY columns into strings (native)
data_n <- data_n %>%
  select(Accepted_name, MAJORITY.x) %>% 
  group_by(Accepted_name) %>%
  mutate(native_climates = paste(MAJORITY.x, collapse = ","))

data_n <- distinct(data_n,
                  Accepted_name,
                  .keep_all = TRUE)
data_n <- data_n[,-c(2)]

#Full climate lists
name_status <- data[,c(1,2)]
full_climates <- left_join(name_status, data_n)
full_climates <- left_join(full_climates, data_e)
full_climates <- left_join(full_climates, data_i)
rm(name_status)

#Add commas to make elements easier to search
full_climates$established_climates<- paste(",",full_climates$established_climates, sep="")
full_climates$invaded_climates<- paste(",",full_climates$invaded_climates, sep="")
full_climates$native_climates<- paste(",",full_climates$native_climates, sep="")

full_climates$established_climates<- paste(full_climates$established_climates,",", sep="")
full_climates$invaded_climates<- paste(full_climates$invaded_climates,",", sep="")
full_climates$native_climates<- paste(full_climates$native_climates,",", sep="")

#Remove NAs
full_climates$native_climates <- gsub("NA", "", full_climates$native_climates)
full_climates$established_climates <- gsub("NA", "", full_climates$established_climates)
full_climates$invaded_climates <- gsub("NA", "", full_climates$invaded_climates)

rm(data_e)
rm(data_i)
rm(data_n)
```

From here, you can either move right into the next section. Or you can skip directly to section 4. Climate Flows Results.

## 3. Recipient Climate Results

Making data table for results

```{r}
#Make a Data Table to Store Destination Results
ClimatesAsDestinations <- data.table()
ClimatesAsDestinations$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
ClimatesAsDestinations$Climate <- c("Tropical", "Arid", "Temperate", "Continental", "Polar")
ClimatesAsDestinations$EstablishedSpecies <- ""
ClimatesAsDestinations$InvasiveSpecies <- ""
```

Count established and invasive species arriving in each climate

```{r}
a <- c(1:5)
for (i in a) {
  
#Need species with invaded climates = whatever Climate is in Row i
Climate_I <- full_climates[grep(paste(ClimatesAsDestinations[i,1] ), full_climates$invaded_climates),]

#Remove excess columns
Climate_I <- Climate_I[,c(1)]

#Label as invasive
Climate_I$type = "1"

#Need species with Established climates = whatever  Climate is in Row i
Climate_E <- full_climates[grep(paste(ClimatesAsDestinations[i,1] ), full_climates$established_climates),]

#Remove invasive species from our established list
Climate_E <-  Climate_E[-(grep(paste(ClimatesAsDestinations[i,1] ), Climate_E$invaded_climates)),]

#Remove excess columns
Climate_E <- Climate_E[,c(1)]

#Label as established
Climate_E$type = "0"

#Rbind Invaders to Established 
Climate <- rbind(Climate_I, Climate_E)

Climate_count <- table(Climate$type)

ClimatesAsDestinations[i,3] <- paste(Climate_count[1])
ClimatesAsDestinations[i,4] <- paste(Climate_count[2])

rm(Climate)
rm(Climate_E)
rm(Climate_I)
rm(Climate_count)
  
 message('Running Climate ', i, ' of 5')

}

#Calculate Invasion Rates
ClimatesAsDestinations$EstablishedSpecies <- as.integer(ClimatesAsDestinations$EstablishedSpecies)
ClimatesAsDestinations$InvasiveSpecies <- as.integer(ClimatesAsDestinations$InvasiveSpecies)
ClimatesAsDestinations$Total <- (ClimatesAsDestinations$EstablishedSpecies + ClimatesAsDestinations$InvasiveSpecies)
ClimatesAsDestinations$InvasionRate <- (ClimatesAsDestinations$InvasiveSpecies/ClimatesAsDestinations$Total)

#Run this - or not - depending on whether you removed islands earlier
#NoI_ClimatesAsDestinations <- ClimatesAsDestinations
NoM_ClimatesAsDestinations <- ClimatesAsDestinations

```

Plotting Invasion Rates for Recipient Climates (Point Plot with Error Bars)

```{r}
#Assign results to correct categories
NoI_ClimatesAsDestinations$IslandData <- paste("Mainlands Only")
new_row    <- data.table(`Climate` = "All")
NOIC <- rbind(NoI_ClimatesAsDestinations, new_row, fill=TRUE) 
NOIC[6,7] <- paste("Mainlands Only")
NOIC[6,5] <- as.numeric(paste(10140))
NOIC[6,4] <- as.numeric(paste(1517))
NOIC[6,6] <- NOIC[6,4] / NOIC[6,5]

NoM_ClimatesAsDestinations$IslandData <- paste("Islands Only")
NOMC <- rbind(NoM_ClimatesAsDestinations, new_row, fill=TRUE)
NOMC[6,7] <- paste("Islands Only")
NOMC[6,5] <- as.numeric(paste(6921))
NOMC[6,4] <- as.numeric(paste(1295))
NOMC[6,6] <- NOMC[6,4] / NOMC[6,5]

ClimatesAsDestinations$IslandData <- paste("All Data")
CAD <- rbind(ClimatesAsDestinations, new_row, fill=TRUE)
CAD[6,7] <- paste("All Data")
CAD[6,5] <- as.numeric(paste(12126))
CAD[6,4] <- as.numeric(paste(2039))
CAD[6,6] <- CAD[6,4] / CAD[6,5]

#Combine results from island data and no island data
pd <- rbind(CAD, NOMC, NOIC)
pd$CIMax <- ""
pd$CIMin <- ""

#Calculate Confidence Intervals
a <- c(1:18)
for (i in a) {
conf <- scoreci(as.integer(paste(pd[i,4])), as.integer(paste(pd[i,5])), conf.level = 0.95)
pd[i,8] <- conf[1]$conf.int[2]
pd[i,9] <- conf[1]$conf.int[1]
}

pd$CIMax <- as.numeric(pd$CIMax)
pd$CIMin <- as.numeric(pd$CIMin)

pd$label <- paste(c("Tropical",
                      "Arid",
                      "Temperate",
                      "Continental",
                      "Polar",
                      "All",
                      "Tropical_IO",
                      "Arid_IO",
                      "Temperate_IO",
                      "Continental_IO",
                      "Polar_IO",
                      "All_IO",
                      "Tropical_MO",
                      "Arid_MO",
                      "Temperate_MO",
                      "Continental_MO",
                      "Polar_MO",
                      "All_MO"))

pd$label <- factor(pd$label, levels=unique(pd$label))

#Build Plot
g <- ggplot(pd, aes(x= fct_inorder(`label`), y=`InvasionRate`, colour=`Climate`))+ 
                stat_summary(geom="point", fun=mean, size= 3.1 ) +
                ylab("Invasion Rate (%)") +
                coord_cartesian(ylim=c(0, 0.34),xlim = c(0.5,18.5), expand = FALSE, clip = "off") +
                theme(legend.position = "none",
                      plot.margin = unit(c(1, 1, 4.5, 1), "lines"),
                      axis.title.x = element_blank(),
                      axis.text.x = element_blank(),
                      axis.title.y = element_text(face="bold")) +
                geom_hline(yintercept= 0.1, linetype="dashed", color = "black", linewidth=0.6) +
                geom_vline(aes(xintercept = 6.5)) +
                geom_vline(aes(xintercept = 12.5)) +
                annotate(geom = "text", x = seq_len(nrow(pd)), y = -0.02 + rep(c(0,-0.015,0,-0.015,0,-0.015), 3),
                         label = pd$Climate, size = 3) +
                annotate(geom = "text", x = 3.5 + (6 * (0:2)), y = -0.06, label = unique(pd$IslandData), 
                         size = 4, fontface = "bold")
p <- g + 
  geom_errorbar(aes(ymin = pd$CIMin, ymax = pd$CIMax), width = 0.2, linewidth =1.2) +
  scale_color_manual(values=c("black", "#ffb703", "#027fdc", "#07c4c5", "#f76f73", "#0218a2",
                             "black", "#ffb703", "#027fdc", "#07c4c5", "#f76f73", "#0218a2",
                             "black", "#ffb703", "#027fdc", "#07c4c5", "#f76f73", "#0218a2"))

```

Significance Testing

```{r}
#Are the proportions for each climate significantly different than each other?
pairwise.prop.test(pd$InvasiveSpecies, pd$Total, conf.level=0.95)
```

## 4. Climate Flows Results

Data Tables for Climate Flows

```{r}
#Start by making data tables for results (change these names if you're using "no islands" or "no mainlands" data)

#First one will store proportions of Established/Invasive
CFs <- data.table()
CFs$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs$From_Tropical<- ""
CFs$From_Arid<- ""
CFs$From_Temperate<- ""
CFs$From_Continental<- ""
CFs$From_Polar<- ""

#Next one will store absolute numbers of established species
CFs_E <- data.table()
CFs_E$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_E$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_E$From_Tropical<- ""
CFs_E$From_Arid<- ""
CFs_E$From_Temperate<- ""
CFs_E$From_Continental<- ""
CFs_E$From_Polar<- ""

#Last one will store absolute numbers of invasive species
CFs_I <- data.table()
CFs_I$Nums <- c(",1,|,2,|,3,|,4,",
                                 ",5,|,6,|,7,|,8,",
                                 ",9,|,10,|,11,|,12,|,13,|,14,|,15,|,16,|,17,",
                                 ",18,|,19,|,20,|,21,|,22,|,23,|,24,|,25,|,26,|,27,|,28,|,29,",
                                 ",30,|,31,")
CFs_I$Climate <- c("To Tropical", "To Arid", "To Temperate", "To Continental", "To Polar")
CFs_I$From_Tropical<- ""
CFs_I$From_Arid<- ""
CFs_I$From_Temperate<- ""
CFs_I$From_Continental<- ""
CFs_I$From_Polar<- ""
```

Climate Flows (Sources to Destinations)

```{r}
a <- c(1:5)
b <- c(1:5)
for (i in a) {

#Pull out species native to ith climate  
Natives <- full_climates[grep(paste(CFs[i,1]), full_climates$native_climates),]
  
for (j in b){
#Pull out established species in ith climate (which are already native to ith climate)
e_count <- nrow(Natives[grep(paste(CFs[j,1]), Natives$established_climates),])

#Pull out invasive species in ith climate (which are already native to ith climate)
i_count <- nrow(Natives[grep(paste(CFs[j,1]), Natives$invaded_climates),])

#Do a little arithmetic
CFs[j,i+2] <- as.numeric(paste(i_count/e_count))
CFs_E[j,i+2] <- as.numeric(paste(e_count))
CFs_I[j,i+2] <- as.numeric(paste(i_count))
  
}
}
rm(Natives)

#Run these three lines if you removed islands above
#CFs_NoI<- CFs
#CFs_E_NoI<- CFs_E
#CFs_I_NoI<- CFs_I

#Run these three lines if you removed mainlands above
#CFs_NoM<- CFs
#CFs_E_NoM<- CFs_E
#CFs_I_NoM<- CFs_I

#Leave data tables with current names if you didn't remove anything. 
#But, if you do all three versions, you have to do this one last. 
```

Plotting Climate Flows Results (Tile Plot)

```{r}
pcfd <- CFs[,-c(1)]
pcfd$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))
pcfd$From_Tropical <- as.numeric(pcfd$From_Tropical)
pcfd$From_Arid <- as.numeric(pcfd$From_Arid)
pcfd$From_Temperate <- as.numeric(pcfd$From_Temperate)
pcfd$From_Continental <- as.numeric(pcfd$From_Continental)
pcfd$From_Polar <- as.numeric(pcfd$From_Polar)
pcfd <- pcfd %>% mutate(across(where(is.numeric), round, digits=2))
mpcfd <- melt(pcfd)

m <- ggplot(mpcfd, aes(x=Climate, y=variable)) +
        geom_tile(aes(fill=value, color=as.factor(Climate), width=0.7, height=0.7), linewidth=2) +
        scale_fill_gradientn(colours=c("white", "red"), name = "Invasion Rate (%)") +
        geom_text(aes(label=value))+
        ylab("Source Climate") +
        xlab("Recipient Climate") +
        theme(axis.title.y = element_text(face="bold"),
              axis.title.x = element_text(face="bold"),
              legend.title = element_text(face="bold"))

mm <- m + scale_color_manual(values=c("#0218a2", "#ffb703", "#f76f73", "#027fdc", "#07c4c5"), guide = 'none') +
    scale_y_discrete(labels=c('Tropical', 'Arid', 'Temperate', 'Continental', 'Polar'))
```

Significance Testing for Climate Flows

```{r}
#Add absolute values to mpcfd
CFs_E <- CFs_E[,-c(1)]
CFs_I <- CFs_I[,-c(1)]

CFs_E$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))
CFs_I$Climate <- factor(c("Tropical", "Arid", "Temperate", "Continental", "Polar"),
                      levels=c("Tropical", "Arid", "Temperate", "Continental", "Polar"))

CFs_E$From_Tropical <- as.numeric(CFs_E$From_Tropical)
CFs_E$From_Arid <- as.numeric(CFs_E$From_Arid)
CFs_E$From_Temperate <- as.numeric(CFs_E$From_Temperate)
CFs_E$From_Continental <- as.numeric(CFs_E$From_Continental)
CFs_E$From_Polar <- as.numeric(CFs_E$From_Polar)

CFs_I$From_Tropical <- as.numeric(CFs_I$From_Tropical)
CFs_I$From_Arid <- as.numeric(CFs_I$From_Arid)
CFs_I$From_Temperate <- as.numeric(CFs_I$From_Temperate)
CFs_I$From_Continental <- as.numeric(CFs_I$From_Continental)
CFs_I$From_Polar <- as.numeric(CFs_I$From_Polar)

m_CFs_E <- melt(CFs_E)
m_CFs_I <- melt(CFs_I)

st <- left_join(mpcfd, m_CFs_E, by=c("Climate"="Climate", "variable"="variable"))
st <- rename(st, "Total_Species" ="value.y")

st <- left_join(st, m_CFs_I, by=c("Climate"="Climate", "variable"="variable"))
st<- rename(st, "Invasive_Species" ="value")

#Are the proportions for each climate significantly different than each other?
pairwise.prop.test(st$Invasive_Species, st$Total_Species, conf.level=0.95)
```
