---
title: "Table1"
author: "Will Pfadenhauer"
date: "2023-06-13"
output: html_document
---

1. Load data & packages
```{r}
require(data.table)
require(dplyr)
require(tidyr)

#Load data
L3_clim <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L3_Better_Climate_data.csv")
L4_clim <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Climate_Data/L4_Better_Climate_data.csv")
```

2. Making L3 island list
```{r}
glonaf_tdwg <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Species_Lists/GloNAF/Region_GloNAF_vanKleunenetal2018Ecology.csv")

extra_islands<- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Extra_Islands.csv")

glonaf_tdwg <- glonaf_tdwg[,-c(1:3,5,7:13)]
glonaf_tdwg <- glonaf_tdwg[grep("-OO", glonaf_tdwg$tdwg4), ]
glonaf_tdwg <- distinct(glonaf_tdwg, 
                        tdwg4,
                        .keep_all = TRUE)
extra_islands <- extra_islands[grep("-OO", extra_islands$tdwg4), ]
all_islands <- rbind(glonaf_tdwg, extra_islands)
all_islands <- distinct(all_islands,
                        tdwg4,
                        .keep_all = TRUE)

#Append climates to L3 island list
check <- left_join(all_islands, L3_clim, by=c("tdwg4"="LEVEL3_COD"))

#Remove column that will be a duplicate once we load in more island data
check <- check[,-c(1)]

#Need to append continental/oceanic island status
island_geos <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Islands_WITH_oceanic_continental_classes.csv")

check <- left_join(check, island_geos, by=c("tdwg4"="tdwg4"))
```


2A. L3 ALL Islands 
```{r}
step2isl_all <- check[grep("1", check$island), ]
  
#2A Tropical island list
l3_trop_isl_all <- step2isl_all[(step2isl_all$MAJORITY < 4.5), ]
clipr::write_clip(l3_trop_isl_all$tdwg4)     

#2B Arid island list
l3_arid_isl_all <- step2isl_all[(step2isl_all$MAJORITY > 4.5 & step2isl_all$MAJORITY < 8.5 ), ]
clipr::write_clip(l3_arid_isl_all$tdwg4)   

#2C Temperate island list
l3_temp_isl_all <- step2isl_all[(step2isl_all$MAJORITY > 8.5 & step2isl_all$MAJORITY < 17.5 ), ]
clipr::write_clip(l3_temp_isl_all$tdwg4) 

#2D Continental island list
l3_cont_isl_all <- step2isl_all[(step2isl_all$MAJORITY > 17.5 & step2isl_all$MAJORITY < 29.5 ), ]
clipr::write_clip(l3_cont_isl_all$tdwg4) 

#2E Polar island list
l3_pol_isl_all <- step2isl_all[(step2isl_all$MAJORITY > 29.5), ]
clipr::write_clip(l3_pol_isl_all$tdwg4) 

```

2B. L3 OCEANIC Islands
```{r}
step2isl_oce <- step2isl_all[grep("Oceanic", step2isl_all$geo), ]
  
#2A Tropical oceanic island list
l3_trop_isl_oce <- step2isl_oce[(step2isl_oce$MAJORITY < 4.5), ]
clipr::write_clip(l3_trop_isl_oce$tdwg4)     

#2B Arid oceanic island list
l3_arid_isl_oce <- step2isl_oce[(step2isl_oce$MAJORITY > 4.5 & step2isl_oce$MAJORITY < 8.5 ), ]
clipr::write_clip(l3_arid_isl_oce$tdwg4)   

#2C Temperate oceanic island list
l3_temp_isl_oce <- step2isl_oce[(step2isl_oce$MAJORITY > 8.5 & step2isl_oce$MAJORITY < 17.5 ), ]
clipr::write_clip(l3_temp_isl_oce$tdwg4) 

#2D Continental oceanic island list
l3_cont_isl_oce <- step2isl_oce[(step2isl_oce$MAJORITY > 17.5 & step2isl_oce$MAJORITY < 29.5 ), ]
clipr::write_clip(l3_cont_isl_oce$tdwg4) 

#2E Polar oceanic island list
l3_pol_isl_oce <- step2isl_oce[(step2isl_oce$MAJORITY > 29.5), ]
clipr::write_clip(l3_pol_isl_oce$tdwg4) 
```

2C. L3 CONTINENTAL Islands
```{r}
step2isl_cont <- step2isl_all[grep("Continental", step2isl_all$geo), ]
  
#2A Tropical continental island list
l3_trop_isl_cont <- step2isl_cont[(step2isl_cont$MAJORITY < 4.5), ]
clipr::write_clip(l3_trop_isl_cont$tdwg4)     

#2B Arid continental island list
l3_arid_isl_cont <- step2isl_cont[(step2isl_cont$MAJORITY > 4.5 & step2isl_cont$MAJORITY < 8.5 ), ]
clipr::write_clip(l3_arid_isl_cont$tdwg4)   

#2C Temperate continental island list
l3_temp_isl_cont <- step2isl_cont[(step2isl_cont$MAJORITY > 8.5 & step2isl_cont$MAJORITY < 17.5 ), ]
clipr::write_clip(l3_temp_isl_cont$tdwg4) 

#2D Continental continental island list
l3_cont_isl_cont <- step2isl_cont[(step2isl_cont$MAJORITY > 17.5 & step2isl_cont$MAJORITY < 29.5 ), ]
clipr::write_clip(l3_cont_isl_cont$tdwg4) 

#2E Polar continental island list
l3_pol_isl_cont <- step2isl_cont[(step2isl_cont$MAJORITY > 29.5), ]
clipr::write_clip(l3_pol_isl_cont$tdwg4) 
```


2D. L3 Mainlands 
```{r}
step2main <- check[grep("0", check$island), ]
  
#2A Tropical mainland list
l3_trop_main <- step2main[(step2main$MAJORITY < 4.5), ]
clipr::write_clip(l3_trop_main$tdwg4)     

#2B Arid mainland list
l3_arid_main <- step2main[(step2main$MAJORITY > 4.5 & step2main$MAJORITY < 8.5 ), ]
clipr::write_clip(l3_arid_main$tdwg4)   

#2C Temperate mainland list
l3_temp_main <- step2main[(step2main$MAJORITY > 8.5 & step2main$MAJORITY < 17.5 ), ]
clipr::write_clip(l3_temp_main$tdwg4) 

#2D Continental mainland list
l3_cont_main <- step2main[(step2main$MAJORITY > 17.5 & step2main$MAJORITY < 29.5 ), ]
clipr::write_clip(l3_cont_main$tdwg4) 

#2E Polar mainland list
l3_pol_main <- step2main[(step2main$MAJORITY > 29.5), ]
clipr::write_clip(l3_pol_main$tdwg4) 

```

3. Making L4 island list
```{r}
glonaf_tdwg <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Species_Lists/GloNAF/Region_GloNAF_vanKleunenetal2018Ecology.csv")

extra_islands<- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Extra_Islands.csv")

glonaf_tdwg <- glonaf_tdwg[,-c(1:3,5,7:13)]
glonaf_tdwg <- glonaf_tdwg[-grep("-OO", glonaf_tdwg$tdwg4), ]
glonaf_tdwg <- distinct(glonaf_tdwg, 
                        tdwg4,
                        .keep_all = TRUE)
extra_islands <- extra_islands[-grep("-OO", extra_islands$tdwg4), ]
all_islands <- rbind(glonaf_tdwg, extra_islands)
all_islands <- distinct(all_islands,
                        tdwg4,
                        .keep_all = TRUE)
all_islands <- all_islands[-grep("000", all_islands$tdwg4), ]

#Append climates to L3 island list
check <- left_join(all_islands, L4_clim, by=c("tdwg4"="Level4_cod"))

#Remove column that will be a duplicate once we load in more island data
check <- check[,-c(1)]

#Need to append continental/oceanic island status
island_geos <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Island_Data/Islands_WITH_oceanic_continental_classes.csv")

check <- left_join(check, island_geos, by=c("tdwg4"="tdwg4"))

```


3A. L4 ALL Islands
```{r}
step3isl_all <- check[grep("1", check$island), ]
  
#2A Tropical Island list
l4_trop_isl_all <- step3isl_all[(step3isl_all$MAJORITY < 4.5), ]
clipr::write_clip(l4_trop_isl_all$tdwg4)     

#2B Arid Island list
l4_arid_isl_all <- step3isl_all[(step3isl_all$MAJORITY > 4.5 & step3isl_all$MAJORITY < 8.5 ), ]
clipr::write_clip(l4_arid_isl_all$tdwg4)   

#2C Temperate Island list
l4_temp_isl_all <- step3isl_all[(step3isl_all$MAJORITY > 8.5 & step3isl_all$MAJORITY < 17.5 ), ]
clipr::write_clip(l4_temp_isl_all$tdwg4) 

#2D Continental Island list
l4_cont_isl_all <- step3isl_all[(step3isl_all$MAJORITY > 17.5 & step3isl_all$MAJORITY < 29.5 ), ]
clipr::write_clip(l4_cont_isl_all$tdwg4) 

#2E Polar Island list
l4_pol_isl_all <- step3isl_all[(step3isl_all$MAJORITY > 29.5), ]
clipr::write_clip(l4_pol_isl_all$tdwg4) 

```

3B. L4 OCEANIC Islands
```{r}
step3isl_oce <- step3isl_all[grep("Oceanic", step3isl_all$geo), ]
  
#3A Tropical oceanic island list
l4_trop_isl_oce <- step3isl_oce[(step3isl_oce$MAJORITY < 4.5), ]
clipr::write_clip(l4_trop_isl_oce$tdwg4)     

#3B Arid oceanic island list
l4_arid_isl_oce <- step3isl_oce[(step3isl_oce$MAJORITY > 4.5 & step3isl_oce$MAJORITY < 8.5 ), ]
clipr::write_clip(l4_arid_isl_oce$tdwg4)   

#3C Temperate oceanic island list
l4_temp_isl_oce <- step3isl_oce[(step3isl_oce$MAJORITY > 8.5 & step3isl_oce$MAJORITY < 17.5 ), ]
clipr::write_clip(l4_temp_isl_oce$tdwg4) 

#3D Continental oceanic island list
l4_cont_isl_oce <- step3isl_oce[(step3isl_oce$MAJORITY > 17.5 & step3isl_oce$MAJORITY < 29.5 ), ]
clipr::write_clip(l4_cont_isl_oce$tdwg4) 

#3E Polar oceanic island list
l4_pol_isl_oce <- step3isl_oce[(step3isl_oce$MAJORITY > 29.5), ]
clipr::write_clip(l4_pol_isl_oce$tdwg4) 
```

3C. L4 CONTINENTAL Islands
```{r}
step3isl_cont <- step3isl_all[grep("Continental", step3isl_all$geo), ]
  
#3A Tropical oceanic island list
l4_trop_isl_cont <- step3isl_cont[(step3isl_cont$MAJORITY < 4.5), ]
clipr::write_clip(l4_trop_isl_cont$tdwg4)     

#3B Arid oceanic island list
l4_arid_isl_cont <- step3isl_cont[(step3isl_cont$MAJORITY > 4.5 & step3isl_cont$MAJORITY < 8.5 ), ]
clipr::write_clip(l4_arid_isl_cont$tdwg4)   

#3C Temperate oceanic island list
l4_temp_isl_cont <- step3isl_cont[(step3isl_cont$MAJORITY > 8.5 & step3isl_cont$MAJORITY < 17.5 ), ]
clipr::write_clip(l4_temp_isl_cont$tdwg4) 

#3D Continental oceanic island list
l4_cont_isl_cont <- step3isl_cont[(step3isl_cont$MAJORITY > 17.5 & step3isl_cont$MAJORITY < 29.5 ), ]
clipr::write_clip(l4_cont_isl_cont$tdwg4) 

#3E Polar oceanic island list
l4_pol_isl_cont <- step3isl_cont[(step3isl_cont$MAJORITY > 29.5), ]
clipr::write_clip(l4_pol_isl_cont$tdwg4) 
```

3D. L4 Mainlands
```{r}
step3main <- check[grep("0", check$island), ]
  
#2A Tropical mainland list
l4_trop_main <- step3main[(step3main$MAJORITY < 4.5), ]
clipr::write_clip(l4_trop_main$tdwg4)     

#2B Arid mainland list
l4_arid_main <- step3main[(step3main$MAJORITY > 4.5 & step3main$MAJORITY < 8.5 ), ]
clipr::write_clip(l4_arid_main$tdwg4)   

#2C Temperate mainland list
l4_temp_main <- step3main[(step3main$MAJORITY > 8.5 & step3main$MAJORITY < 17.5 ), ]
clipr::write_clip(l4_temp_main$tdwg4) 

#2D Continental mainland list
l4_cont_main <- step3main[(step3main$MAJORITY > 17.5 & step3main$MAJORITY < 29.5 ), ]
clipr::write_clip(l4_cont_main$tdwg4) 

#2E Polar mainland list
l4_pol_main <- step3main[(step3main$MAJORITY > 29.5), ]
clipr::write_clip(l4_pol_main$tdwg4) 
```


4. Actually calculating the values for Table 1. Can skip directly to the next step if you're just re-running code. 
```{r}
# data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/December_Database_Backup.csv")
# 
# #First, remove NA regions
# NA_regions <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/NA_regions_codes.csv")
# 
# #L2 Established
# a <- c(1:1)
# for (i in a) {
# data$Established_L2 <- gsub(paste(NA_regions[i,1]), ",", data$Established_L2)
# message('Running region ', i, ' of 1')
# }
# 
# #L2 Invasive
# for (i in a) {
# data$Invasive_L2 <- gsub(paste(NA_regions[i,1]), ",", data$Invasive_L2)
# message('Running region ', i, ' of 1')
# }
# 
# #L3 Established
# b <- c(1:117)
# for (i in b) {
# data$Established_L3 <- gsub(paste(NA_regions[i,2]), ",", data$Established_L3)
# message('Running region ', i, ' of 117')
# }
# 
# #L3 Invasive
# for (i in b) {
# data$Invasive_L3 <- gsub(paste(NA_regions[i,2]), ",", data$Invasive_L3)
# message('Running region ', i, ' of 117')
# }
# 
# #L4 Established
# c <- c(1:141)
# for (i in c) {
# data$Established_L4 <- gsub(paste(NA_regions[i,3]), ",", data$Established_L4)
# message('Running region ', i, ' of 141')
# }
# 
# #L4 Invasive
# for (i in c) {
# data$Invasive_L4 <- gsub(paste(NA_regions[i,3]), ",", data$Invasive_L4)
# message('Running region ', i, ' of 141')
# }
# 
# fwrite(data, "C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Dec_Dat_Bac_WO_NA_regions.csv")
```


4. (preface) Bring in codes and format them
```{r}
data <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Dec_Dat_Bac_WO_NA_regions.csv")
codes <- fread("C:/Users/wpfadenhauer/OneDrive - University of Massachusetts/CH2_InvaderFlows/R_files/Table1_Codes_R.csv")

codes$Codes <- gsub(",", ",|,", codes$Codes)
codes$Codes <- paste(",",codes$Codes, sep = "")
codes$Codes <- paste(codes$Codes,",", sep = "")
codes$Codes <- gsub(",", "\\b", codes$Codes, fixed = TRUE)

results <- codes
results$Codes <- ""
Prop_results <- results
Count_results <- results
Reg_Count_results <- results
Sp_Count_results <- results
rm(results)

#Define function to remove duplicates
remove_duplicates <- function(col1, col2) {

  list1 <- strsplit(col1, ",")
  list2 <- strsplit(col2, ",")
  
  result <- Map(function(x, y) paste(setdiff(y, x), collapse = ","), list1, list2)
  
  unlist(result)
}
```

4A. L2 data prep
```{r}
data$Established_L2 <- remove_duplicates(data$Native_L2, data$Established_L2)
data$Invasive_L2 <- remove_duplicates(data$Native_L2, data$Invasive_L2)

data_l2 <- data
```


4B. L2 For-Loop
```{r}

est_sp_l2 <- list() 
inv_sp_l2 <- list()
exp <- list()
regions <- list()

a <- c(1:25)
for (i in a) {
  
est_sp_l2[[i]] <- data_l2[grep(paste(codes[i,5]), data_l2$Established_L2), ]
inv_sp_l2[[i]] <- data_l2[grep(paste(codes[i,5]), data_l2$Invasive_L2), ]

exp[[i]] <- separate_rows(est_sp_l2[[i]], Established_L2, sep = ",")

exp[[i]] <- exp[[i]][grep(paste(codes[i,5]), exp[[i]]$Established_L2), ]

regions[[i]] <- distinct(exp[[i]],
                        Established_L2,
                        .keep_all = TRUE)

Reg_Count_results[i,5] <- nrow(regions[[i]])


est_sp_l2[[i]] <- distinct(est_sp_l2[[i]],
                        Accepted_name,
                        .keep_all = TRUE)

Sp_Count_results[i,5] <- nrow(est_sp_l2[[i]])

inv_sp_l2[[i]] <- distinct(inv_sp_l2[[i]],
                        Accepted_name,
                        .keep_all = TRUE)

Prop_results[i,5] <-  nrow(inv_sp_l2[[i]])/nrow(est_sp_l2[[i]])
}
```


4C. L3 data prep
```{r}
data$Established_L3 <- remove_duplicates(data$Native_L3, data$Established_L3)
data$Invasive_L3 <- remove_duplicates(data$Native_L3, data$Invasive_L3)

data_l3 <- data
```

4D. L3 For-loop
```{r}

est_sp_l3 <- list() 
inv_sp_l3 <- list()
exp <- list()
regions <- list()

a <- c(26:50)
for (i in a) {
  
est_sp_l3[[i]] <- data_l3[grep(paste(codes[i,5]), data_l3$Established_L3), ]
inv_sp_l3[[i]] <- data_l3[grep(paste(codes[i,5]), data_l3$Invasive_L3), ]

exp[[i]] <- separate_rows(est_sp_l3[[i]], Established_L3, sep = ",")

exp[[i]] <- exp[[i]][grep(paste(codes[i,5]), exp[[i]]$Established_L3), ]

regions[[i]] <- distinct(exp[[i]],
                        Established_L3,
                        .keep_all = TRUE)

Reg_Count_results[i,5] <- nrow(regions[[i]])


est_sp_l3[[i]] <- distinct(est_sp_l3[[i]],
                        Accepted_name,
                        .keep_all = TRUE)

Sp_Count_results[i,5] <- nrow(est_sp_l3[[i]])

inv_sp_l3[[i]] <- distinct(inv_sp_l3[[i]],
                        Accepted_name,
                        .keep_all = TRUE)

Prop_results[i,5] <-  nrow(inv_sp_l3[[i]])/nrow(est_sp_l3[[i]])
}

```

4E. L4 data prep
```{r}
data$Established_L4 <- remove_duplicates(data$Native_L4, data$Established_L4)
data$Invasive_L4 <- remove_duplicates(data$Native_L4, data$Invasive_L4)

data_l4 <- data
```


4F. L4 For-Loop
```{r}

est_sp_l4 <- list() 
inv_sp_l4 <- list()
exp <- list()
regions <- list()

a <- c(51:75)
for (i in a) {
  
est_sp_l4[[i]] <- data_l4[grep(paste(codes[i,5]), data_l4$Established_L4), ]
inv_sp_l4[[i]] <- data_l4[grep(paste(codes[i,5]), data_l4$Invasive_L4), ]

exp[[i]] <- separate_rows(est_sp_l4[[i]], Established_L4, sep = ",")

exp[[i]] <- exp[[i]][grep(paste(codes[i,5]), exp[[i]]$Established_L4), ]

regions[[i]] <- distinct(exp[[i]],
                        Established_L4,
                        .keep_all = TRUE)

Reg_Count_results[i,5] <- nrow(regions[[i]])


est_sp_l4[[i]] <- distinct(est_sp_l4[[i]],
                        Accepted_name,
                        .keep_all = TRUE)

Sp_Count_results[i,5] <- nrow(est_sp_l4[[i]])

inv_sp_l4[[i]] <- distinct(inv_sp_l4[[i]],
                        Accepted_name,
                        .keep_all = TRUE)

Prop_results[i,5] <-  nrow(inv_sp_l4[[i]])/nrow(est_sp_l4[[i]])
}

```
